---
title: "First pass analysis"
author: "Nikki Reck"
format: html
editor: visual
---

## Background

This is my first pass analysis for my data on red squirrel vocal distinctiveness. I will be cleaning the data so that it is compatible with R (renaming columns, and creating new columns). Then I will take averages of each call from each animal, then take averages of all calls per animal. Then I will look at how the two populations may or may not differ.

Load in packages and the data sets I will be using.

```{r}
library(tidyverse)

sq_data <- read_csv("data_raw/combined_sq_data4.csv")
view(sq_data)

detections <- read_csv("data_raw/01jan2026_detections4.csv")

# Clean break_s: "_" = missing, then make numeric
detections <- detections %>%
  mutate(
    break_s = as.character(break_s),
    break_s = na_if(break_s, "_"),
    break_s = as.numeric(break_s)
  )

# combine data
sq_rattle_data <- detections %>%
  left_join(sq_data, by = "midden")

```

## Call level Averages

Taking the averages of each of the calls per animal. Fix any oddities.

```{r}

# Call-level averages by sq_ID (one row per call)

call_avgs <- sq_rattle_data %>%
  group_by(sq_ID, filename) %>%
  summarise(
    no_notes = n(),
    call_duration = first(call_duration),
    note_rate = no_notes / call_duration,

    duration = mean(duration, na.rm = TRUE),
    break_s = mean(break_s, na.rm = TRUE),
    disttomax = mean(disttomax, na.rm = TRUE),

    rms = mean(rms, na.rm = TRUE),
    energy = mean(energy, na.rm = TRUE),
    peaktopeak = mean(peaktopeak, na.rm = TRUE),

    peak_freq_mean = mean(peak_freq_mean, na.rm = TRUE),
    peak_ampl_mean = mean(peak_ampl_mean, na.rm = TRUE),
    min_freq_mean = mean(min_freq_mean, na.rm = TRUE),
    max_freq_mean = mean(max_freq_mean, na.rm = TRUE),
    bandw_mean = mean(bandw_mean, na.rm = TRUE),

    quart_25_mean = mean(quart_25_mean, na.rm = TRUE),
    quart_50_mean = mean(quart_50_mean, na.rm = TRUE),
    quart_75_mean = mean(quart_75_mean, na.rm = TRUE),

    entropy_mean = mean(entropy_mean, na.rm = TRUE)
  ) %>%
  ungroup()
call_avgs


```

## Squirrel level averages

Taking the averages of all the calls per animal.

```{r}


sq_avgs <- call_avgs %>%
  group_by(sq_ID) %>%
  summarise(
    n_calls = n(),

    # averages of call-level traits
    call_duration = mean(call_duration, na.rm = TRUE),
    no_notes = mean(no_notes, na.rm = TRUE),
    note_rate = mean(note_rate, na.rm = TRUE),

    duration = mean(duration, na.rm = TRUE),
    break_s = mean(break_s, na.rm = TRUE),
    disttomax = mean(disttomax, na.rm = TRUE),

    rms = mean(rms, na.rm = TRUE),
    energy = mean(energy, na.rm = TRUE),
    peaktopeak = mean(peaktopeak, na.rm = TRUE),

    peak_freq_mean = mean(peak_freq_mean, na.rm = TRUE),
    peak_ampl_mean = mean(peak_ampl_mean, na.rm = TRUE),
    min_freq_mean = mean(min_freq_mean, na.rm = TRUE),
    max_freq_mean = mean(max_freq_mean, na.rm = TRUE),
    bandw_mean = mean(bandw_mean, na.rm = TRUE),

    quart_25_mean = mean(quart_25_mean, na.rm = TRUE),
    quart_50_mean = mean(quart_50_mean, na.rm = TRUE),
    quart_75_mean = mean(quart_75_mean, na.rm = TRUE),

    entropy_mean = mean(entropy_mean, na.rm = TRUE)
  ) %>%
  ungroup()
sq_avgs


# worried that sqs with 8 calls vs sqs with 4-5 calls will create bias. Does having extra calls (like 8 instead of 5) change the squirrel’s average? Gonna check differences to see if i need to standardize sqs to 5 calls and eliminate extra calls. I calculated squirrel-level means two ways.first, I used all available calls per squirrel and using the first five calls per squirrel as a sensitivity check.

# all calls
sq_avgs_all <- call_avgs %>%
  group_by(sq_ID) %>%
  summarise(
    n_calls = n(),

    call_duration = mean(call_duration, na.rm = TRUE),
    no_notes     = mean(no_notes, na.rm = TRUE),
    note_rate    = mean(note_rate, na.rm = TRUE),

    duration     = mean(duration, na.rm = TRUE),
    break_s      = mean(break_s, na.rm = TRUE),
    disttomax    = mean(disttomax, na.rm = TRUE),

    rms          = mean(rms, na.rm = TRUE),
    energy       = mean(energy, na.rm = TRUE),
    peaktopeak   = mean(peaktopeak, na.rm = TRUE),

    peak_freq_mean  = mean(peak_freq_mean, na.rm = TRUE),
    peak_ampl_mean  = mean(peak_ampl_mean, na.rm = TRUE),
    min_freq_mean   = mean(min_freq_mean, na.rm = TRUE),
    max_freq_mean   = mean(max_freq_mean, na.rm = TRUE),
    bandw_mean      = mean(bandw_mean, na.rm = TRUE),

    quart_25_mean = mean(quart_25_mean, na.rm = TRUE),
    quart_50_mean = mean(quart_50_mean, na.rm = TRUE),
    quart_75_mean = mean(quart_75_mean, na.rm = TRUE),

    entropy_mean  = mean(entropy_mean, na.rm = TRUE)
  ) %>%
  ungroup()


# first 5 calls per squirrel
sq_avgs_5 <- call_avgs %>%
  group_by(sq_ID) %>%
  slice_head(n = 5) %>%
  summarise(
    n_calls = n(),

    call_duration = mean(call_duration, na.rm = TRUE),
    no_notes     = mean(no_notes, na.rm = TRUE),
    note_rate    = mean(note_rate, na.rm = TRUE),

    duration     = mean(duration, na.rm = TRUE),
    break_s      = mean(break_s, na.rm = TRUE),
    disttomax    = mean(disttomax, na.rm = TRUE),

    rms          = mean(rms, na.rm = TRUE),
    energy       = mean(energy, na.rm = TRUE),
    peaktopeak   = mean(peaktopeak, na.rm = TRUE),

    peak_freq_mean  = mean(peak_freq_mean, na.rm = TRUE),
    peak_ampl_mean  = mean(peak_ampl_mean, na.rm = TRUE),
    min_freq_mean   = mean(min_freq_mean, na.rm = TRUE),
    max_freq_mean   = mean(max_freq_mean, na.rm = TRUE),
    bandw_mean      = mean(bandw_mean, na.rm = TRUE),

    quart_25_mean = mean(quart_25_mean, na.rm = TRUE),
    quart_50_mean = mean(quart_50_mean, na.rm = TRUE),
    quart_75_mean = mean(quart_75_mean, na.rm = TRUE),

    entropy_mean  = mean(entropy_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# compare parameters
compare_entropy <- sq_avgs_all %>%
  select(sq_ID, entropy_all = entropy_mean) %>%
  left_join(
    sq_avgs_5 %>% select(sq_ID, entropy_5 = entropy_mean),
    by = "sq_ID"
  )

p_entropy <- compare_entropy %>%
  ggplot(aes(x = entropy_5, y = entropy_all)) +
  geom_point(na.rm = TRUE) +
  geom_abline() +
  labs(
    x = "Entropy mean (first 5 calls)",
    y = "Entropy mean (all calls)"
  )

p_entropy

ggsave("figures/compare_entropy.png", plot = p_entropy, width = 7, height = 5, dpi = 300)

compare_call_duration <- sq_avgs_all %>%
  select(sq_ID, call_duration_all = call_duration) %>%
  left_join(
    sq_avgs_5 %>% select(sq_ID, call_duration_5 = call_duration),
    by = "sq_ID"
  )
p_call_duration <- compare_call_duration %>%
  ggplot(aes(x = call_duration_5, y = call_duration_all)) +
  geom_point(na.rm = TRUE) +
  geom_abline() +
  labs(
    x = "Call duration mean (first 5 calls)",
    y = "Call duration mean (all calls)"
  )

p_call_duration

ggsave("figures/compare_call_duration.png",
       plot = p_call_duration,
       width = 7, height = 5, dpi = 300)

compare_no_notes <- sq_avgs_all %>%
  select(sq_ID, no_notes_all = no_notes) %>%
  left_join(
    sq_avgs_5 %>% select(sq_ID, no_notes_5 = no_notes),
    by = "sq_ID"
  )

p_no_notes <- compare_no_notes %>%
  ggplot(aes(x = no_notes_5, y = no_notes_all)) +
  geom_point(na.rm = TRUE) +
  geom_abline() +
  labs(
    x = "Mean no. notes (first 5 calls)",
    y = "Mean no. notes (all calls)"
  )

p_no_notes

ggsave("figures/compare_no_notes.png",
       plot = p_no_notes,
       width = 7, height = 5, dpi = 300)

compare_note_rate <- sq_avgs_all %>%
  select(sq_ID, note_rate_all = note_rate) %>%
  left_join(
    sq_avgs_5 %>% select(sq_ID, note_rate_5 = note_rate),
    by = "sq_ID"
  )

p_note_rate <- compare_note_rate %>%
  ggplot(aes(x = note_rate_5, y = note_rate_all)) +
  geom_point(na.rm = TRUE) +
  geom_abline() +
  labs(
    x = "Note rate mean (first 5 calls)",
    y = "Note rate mean (all calls)"
  )

p_note_rate

ggsave("figures/compare_note_rate.png",
       plot = p_note_rate,
       width = 7, height = 5, dpi = 300)


compare_duration <- sq_avgs_all %>%
  select(sq_ID, duration_all = duration) %>%
  left_join(
    sq_avgs_5 %>% select(sq_ID, duration_5 = duration),
    by = "sq_ID"
  )

p_duration <- compare_duration %>%
  ggplot(aes(x = duration_5, y = duration_all)) +
  geom_point(na.rm = TRUE) +
  geom_abline() +
  labs(
    x = "Mean note duration (first 5 calls)",
    y = "Mean note duration (all calls)"
  )

p_duration

ggsave("figures/compare_duration.png",
       plot = p_duration,
       width = 7, height = 5, dpi = 300)


compare_break_s <- sq_avgs_all %>%
  select(sq_ID, break_s_all = break_s) %>%
  left_join(
    sq_avgs_5 %>% select(sq_ID, break_s_5 = break_s),
    by = "sq_ID"
  )

p_break_s <- compare_break_s %>%
  ggplot(aes(x = break_s_5, y = break_s_all)) +
  geom_point(na.rm = TRUE) +
  geom_abline() +
  labs(
    x = "Mean break_s (first 5 calls)",
    y = "Mean break_s (all calls)"
  )

p_break_s

ggsave("figures/compare_break_s.png",
       plot = p_break_s,
       width = 7, height = 5, dpi = 300)

compare_disttomax <- sq_avgs_all %>%
  select(sq_ID, disttomax_all = disttomax) %>%
  left_join(
    sq_avgs_5 %>% select(sq_ID, disttomax_5 = disttomax),
    by = "sq_ID"
  )

p_disttomax <- compare_disttomax %>%
  ggplot(aes(x = disttomax_5, y = disttomax_all)) +
  geom_point(na.rm = TRUE) +
  geom_abline() +
  labs(
    x = "Mean disttomax (first 5 calls)",
    y = "Mean disttomax (all calls)"
  )

p_disttomax

ggsave("figures/compare_disttomax.png",
       plot = p_disttomax,
       width = 7, height = 5, dpi = 300)

compare_rms <- sq_avgs_all %>%
  select(sq_ID, rms_all = rms) %>%
  left_join(
    sq_avgs_5 %>% select(sq_ID, rms_5 = rms),
    by = "sq_ID"
  )

p_rms <- compare_rms %>%
  ggplot(aes(x = rms_5, y = rms_all)) +
  geom_point(na.rm = TRUE) +
  geom_abline() +
  labs(
    x = "Mean RMS (first 5 calls)",
    y = "Mean RMS (all calls)"
  )

p_rms

ggsave("figures/compare_rms.png",
       plot = p_rms,
       width = 7, height = 5, dpi = 300)

compare_energy <- sq_avgs_all %>%
  select(sq_ID, energy_all = energy) %>%
  left_join(
    sq_avgs_5 %>% select(sq_ID, energy_5 = energy),
    by = "sq_ID"
  )

p_energy <- compare_energy %>%
  ggplot(aes(x = energy_5, y = energy_all)) +
  geom_point(na.rm = TRUE) +
  geom_abline() +
  labs(
    x = "Mean energy (first 5 calls)",
    y = "Mean energy (all calls)"
  )

p_energy

ggsave("figures/compare_energy.png",
       plot = p_energy,
       width = 7, height = 5, dpi = 300)

compare_peaktopeak <- sq_avgs_all %>%
  select(sq_ID, peaktopeak_all = peaktopeak) %>%
  left_join(
    sq_avgs_5 %>% select(sq_ID, peaktopeak_5 = peaktopeak),
    by = "sq_ID"
  )
p_peaktopeak <- compare_peaktopeak %>%
  ggplot(aes(x = peaktopeak_5, y = peaktopeak_all)) +
  geom_point(na.rm = TRUE) +
  geom_abline() +
  labs(
    x = "Mean peak-to-peak (first 5 calls)",
    y = "Mean peak-to-peak (all calls)"
  )

p_peaktopeak

ggsave("figures/compare_peaktopeak.png",
       plot = p_peaktopeak,
       width = 7, height = 5, dpi = 300)

compare_peak_freq <- sq_avgs_all %>%
  select(sq_ID, peak_freq_all = peak_freq_mean) %>%
  left_join(
    sq_avgs_5 %>% select(sq_ID, peak_freq_5 = peak_freq_mean),
    by = "sq_ID"
  )

p_peak_freq <- compare_peak_freq %>%
  ggplot(aes(x = peak_freq_5, y = peak_freq_all)) +
  geom_point(na.rm = TRUE) +
  geom_abline() +
  labs(
    x = "Peak frequency mean (first 5 calls)",
    y = "Peak frequency mean (all calls)"
  )

p_peak_freq

ggsave("figures/compare_peak_freq_mean.png",
       plot = p_peak_freq,
       width = 7, height = 5, dpi = 300)

compare_peak_ampl <- sq_avgs_all %>%
  select(sq_ID, peak_ampl_all = peak_ampl_mean) %>%
  left_join(
    sq_avgs_5 %>% select(sq_ID, peak_ampl_5 = peak_ampl_mean),
    by = "sq_ID"
  )

p_peak_ampl <- compare_peak_ampl %>%
  ggplot(aes(x = peak_ampl_5, y = peak_ampl_all)) +
  geom_point(na.rm = TRUE) +
  geom_abline() +
  labs(
    x = "Peak amplitude mean (first 5 calls)",
    y = "Peak amplitude mean (all calls)"
  )

p_peak_ampl

ggsave("figures/compare_peak_ampl_mean.png",
       plot = p_peak_ampl,
       width = 7, height = 5, dpi = 300)

compare_min_freq <- sq_avgs_all %>%
  select(sq_ID, min_freq_all = min_freq_mean) %>%
  left_join(
    sq_avgs_5 %>% select(sq_ID, min_freq_5 = min_freq_mean),
    by = "sq_ID"
  )
p_min_freq <- compare_min_freq %>%
  ggplot(aes(x = min_freq_5, y = min_freq_all)) +
  geom_point(na.rm = TRUE) +
  geom_abline() +
  labs(
    x = "Min frequency mean (first 5 calls)",
    y = "Min frequency mean (all calls)"
  )

p_min_freq

ggsave("figures/compare_min_freq_mean.png",
       plot = p_min_freq,
       width = 7, height = 5, dpi = 300)

compare_max_freq <- sq_avgs_all %>%
  select(sq_ID, max_freq_all = max_freq_mean) %>%
  left_join(
    sq_avgs_5 %>% select(sq_ID, max_freq_5 = max_freq_mean),
    by = "sq_ID"
  )

p_max_freq <- compare_max_freq %>%
  ggplot(aes(x = max_freq_5, y = max_freq_all)) +
  geom_point(na.rm = TRUE) +
  geom_abline() +
  labs(
    x = "Max frequency mean (first 5 calls)",
    y = "Max frequency mean (all calls)"
  )

p_max_freq

ggsave("figures/compare_max_freq_mean.png",
       plot = p_max_freq,
       width = 7, height = 5, dpi = 300)

compare_bandw <- sq_avgs_all %>%
  select(sq_ID, bandw_all = bandw_mean) %>%
  left_join(
    sq_avgs_5 %>% select(sq_ID, bandw_5 = bandw_mean),
    by = "sq_ID"
  )
p_bandw <- compare_bandw %>%
  ggplot(aes(x = bandw_5, y = bandw_all)) +
  geom_point(na.rm = TRUE) +
  geom_abline() +
  labs(
    x = "Bandwidth mean (first 5 calls)",
    y = "Bandwidth mean (all calls)"
  )

p_bandw

ggsave("figures/compare_bandw_mean.png",
       plot = p_bandw,
       width = 7, height = 5, dpi = 300)

compare_quart25 <- sq_avgs_all %>%
  select(sq_ID, quart25_all = quart_25_mean) %>%
  left_join(
    sq_avgs_5 %>% select(sq_ID, quart25_5 = quart_25_mean),
    by = "sq_ID"
  )
p_quart25 <- compare_quart25 %>%
  ggplot(aes(x = quart25_5, y = quart25_all)) +
  geom_point(na.rm = TRUE) +
  geom_abline() +
  labs(
    x = "Quartile 25% mean (first 5 calls)",
    y = "Quartile 25% mean (all calls)"
  )

p_quart25

ggsave("figures/compare_quart_25_mean.png",
       plot = p_quart25,
       width = 7, height = 5, dpi = 300)

compare_quart50 <- sq_avgs_all %>%
  select(sq_ID, quart50_all = quart_50_mean) %>%
  left_join(
    sq_avgs_5 %>% select(sq_ID, quart50_5 = quart_50_mean),
    by = "sq_ID"
  )
p_quart50 <- compare_quart50 %>%
  ggplot(aes(x = quart50_5, y = quart50_all)) +
  geom_point(na.rm = TRUE) +
  geom_abline() +
  labs(
    x = "Quartile 50% mean (first 5 calls)",
    y = "Quartile 50% mean (all calls)"
  )

p_quart50

ggsave("figures/compare_quart_50_mean.png",
       plot = p_quart50,
       width = 7, height = 5, dpi = 300)

compare_quart75 <- sq_avgs_all %>%
  select(sq_ID, quart75_all = quart_75_mean) %>%
  left_join(
    sq_avgs_5 %>% select(sq_ID, quart75_5 = quart_75_mean),
    by = "sq_ID"
  )
p_quart75 <- compare_quart75 %>%
  ggplot(aes(x = quart75_5, y = quart75_all)) +
  geom_point(na.rm = TRUE) +
  geom_abline() +
  labs(
    x = "Quartile 75% mean (first 5 calls)",
    y = "Quartile 75% mean (all calls)"
  )

p_quart75

ggsave("figures/compare_quart_75_mean.png",
       plot = p_quart75,
       width = 7, height = 5, dpi = 300)

# All parameters look pretty good. Going to keep all calls

```

## Visualization for averages- MG

Create figures for a clear look into data.

```{r}

# Just going to look at figs for mt graham squirrels.

# add the site  onto the call data, then keep only Mount Graham calls
call_avgs_mg <- call_avgs %>%
  left_join(
    sq_data_avgs %>% distinct(sq_ID, site),
    by = "sq_ID"
  ) %>%
  filter(site == "MG")


# for MG squirrels, calculate the mean and SD of call_duration across calls
call_duration_by_sq_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_call_duration = mean(call_duration, na.rm = TRUE),
    sd_call_duration = sd(call_duration, na.rm = TRUE)
  ) %>%
  ungroup()

p_call_duration_mg <- call_duration_by_sq_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_call_duration), y = mean_call_duration)) +
  geom_pointrange(
    aes(ymin = mean_call_duration - sd_call_duration,
        ymax = mean_call_duration + sd_call_duration),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Call duration: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_call_duration_mg

ggsave("figures/var_call_duration_MG.png",
       plot = p_call_duration_mg,
       width = 8, height = 5, dpi = 300)

#  for MG squirrels, calculate mean and SD of note_rate across calls
note_rate_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_note_rate = mean(note_rate, na.rm = TRUE),
    sd_note_rate = sd(note_rate, na.rm = TRUE)
  ) %>%
  ungroup()
p_note_rate_mg <- note_rate_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_note_rate), y = mean_note_rate)) +
  geom_pointrange(
    aes(ymin = mean_note_rate - sd_note_rate,
        ymax = mean_note_rate + sd_note_rate),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Note rate: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_note_rate_mg
ggsave("figures/var_note_rate_MG.png",
       plot = p_note_rate_mg,
       width = 8, height = 5, dpi = 300)

# duration param MG
# mean + SD of duration for each MG squirrel
duration_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_duration = mean(duration, na.rm = TRUE),
    sd_duration = sd(duration, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_duration_mg <- duration_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_duration), y = mean_duration)) +
  geom_pointrange(
    aes(ymin = mean_duration - sd_duration,
        ymax = mean_duration + sd_duration),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Duration: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_duration_mg

# save into figures folder
ggsave("figures/var_duration_MG.png",
       plot = p_duration_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of break_s for each MG squirrel
break_s_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_break_s = mean(break_s, na.rm = TRUE),
    sd_break_s = sd(break_s, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_break_s_mg <- break_s_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_break_s), y = mean_break_s)) +
  geom_pointrange(
    aes(ymin = mean_break_s - sd_break_s,
        ymax = mean_break_s + sd_break_s),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "break_s: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_break_s_mg

# save into figures folder
ggsave("figures/var_break_s_MG.png",
       plot = p_break_s_mg,
       width = 8, height = 5, dpi = 300)


# mean + SD of disttomax for each MG squirrel
disttomax_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_disttomax = mean(disttomax, na.rm = TRUE),
    sd_disttomax = sd(disttomax, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_disttomax_mg <- disttomax_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_disttomax), y = mean_disttomax)) +
  geom_pointrange(
    aes(ymin = mean_disttomax - sd_disttomax,
        ymax = mean_disttomax + sd_disttomax),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "disttomax: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_disttomax_mg

# save into figures folder
ggsave("figures/var_disttomax_MG.png",
       plot = p_disttomax_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of rms for each MG squirrel
rms_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_rms = mean(rms, na.rm = TRUE),
    sd_rms = sd(rms, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_rms_mg <- rms_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_rms), y = mean_rms)) +
  geom_pointrange(
    aes(ymin = mean_rms - sd_rms,
        ymax = mean_rms + sd_rms),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "RMS: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_rms_mg

# save into figures folder
ggsave("figures/var_rms_MG.png",
       plot = p_rms_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of energy for each MG squirrel
energy_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_energy = mean(energy, na.rm = TRUE),
    sd_energy = sd(energy, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_energy_mg <- energy_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_energy), y = mean_energy)) +
  geom_pointrange(
    aes(ymin = mean_energy - sd_energy,
        ymax = mean_energy + sd_energy),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Energy: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_energy_mg

# save into figures folder
ggsave("figures/var_energy_MG.png",
       plot = p_energy_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of peaktopeak for each MG squirrel
peaktopeak_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_peaktopeak = mean(peaktopeak, na.rm = TRUE),
    sd_peaktopeak = sd(peaktopeak, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_peaktopeak_mg <- peaktopeak_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_peaktopeak), y = mean_peaktopeak)) +
  geom_pointrange(
    aes(ymin = mean_peaktopeak - sd_peaktopeak,
        ymax = mean_peaktopeak + sd_peaktopeak),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Peak-to-peak: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_peaktopeak_mg

# save into figures folder
ggsave("figures/var_peaktopeak_MG.png",
       plot = p_peaktopeak_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of peak_freq_mean for each MG squirrel
peak_freq_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_peak_freq = mean(peak_freq_mean, na.rm = TRUE),
    sd_peak_freq = sd(peak_freq_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_peak_freq_mg <- peak_freq_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_peak_freq), y = mean_peak_freq)) +
  geom_pointrange(
    aes(ymin = mean_peak_freq - sd_peak_freq,
        ymax = mean_peak_freq + sd_peak_freq),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Peak frequency mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_peak_freq_mg

# save into figures folder
ggsave("figures/var_peak_freq_mean_MG.png",
       plot = p_peak_freq_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of peak_ampl_mean for each MG squirrel
peak_ampl_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_peak_ampl = mean(peak_ampl_mean, na.rm = TRUE),
    sd_peak_ampl = sd(peak_ampl_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_peak_ampl_mg <- peak_ampl_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_peak_ampl), y = mean_peak_ampl)) +
  geom_pointrange(
    aes(ymin = mean_peak_ampl - sd_peak_ampl,
        ymax = mean_peak_ampl + sd_peak_ampl),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Peak amplitude mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_peak_ampl_mg

# save into figures folder
ggsave("figures/var_peak_ampl_mean_MG.png",
       plot = p_peak_ampl_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of min_freq_mean for each MG squirrel
min_freq_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_min_freq = mean(min_freq_mean, na.rm = TRUE),
    sd_min_freq = sd(min_freq_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_min_freq_mg <- min_freq_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_min_freq), y = mean_min_freq)) +
  geom_pointrange(
    aes(ymin = mean_min_freq - sd_min_freq,
        ymax = mean_min_freq + sd_min_freq),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Min frequency mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_min_freq_mg

# save into figures folder
ggsave("figures/var_min_freq_mean_MG.png",
       plot = p_min_freq_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of max_freq_mean for each MG squirrel
max_freq_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_max_freq = mean(max_freq_mean, na.rm = TRUE),
    sd_max_freq = sd(max_freq_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_max_freq_mg <- max_freq_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_max_freq), y = mean_max_freq)) +
  geom_pointrange(
    aes(ymin = mean_max_freq - sd_max_freq,
        ymax = mean_max_freq + sd_max_freq),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Max frequency mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_max_freq_mg

# save into figures folder
ggsave("figures/var_max_freq_mean_MG.png",
       plot = p_max_freq_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of bandw_mean for each MG squirrel
bandw_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_bandw = mean(bandw_mean, na.rm = TRUE),
    sd_bandw = sd(bandw_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_bandw_mg <- bandw_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_bandw), y = mean_bandw)) +
  geom_pointrange(
    aes(ymin = mean_bandw - sd_bandw,
        ymax = mean_bandw + sd_bandw),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Bandwidth mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_bandw_mg

# save into figures folder
ggsave("figures/var_bandw_mean_MG.png",
       plot = p_bandw_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of quart_25_mean for each MG squirrel
quart25_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_quart25 = mean(quart_25_mean, na.rm = TRUE),
    sd_quart25 = sd(quart_25_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_quart25_mg <- quart25_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_quart25), y = mean_quart25)) +
  geom_pointrange(
    aes(ymin = mean_quart25 - sd_quart25,
        ymax = mean_quart25 + sd_quart25),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Quartile 25% mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_quart25_mg

# save into figures folder
ggsave("figures/var_quart_25_mean_MG.png",
       plot = p_quart25_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of quart_50_mean for each MG squirrel
quart50_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_quart50 = mean(quart_50_mean, na.rm = TRUE),
    sd_quart50 = sd(quart_50_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_quart50_mg <- quart50_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_quart50), y = mean_quart50)) +
  geom_pointrange(
    aes(ymin = mean_quart50 - sd_quart50,
        ymax = mean_quart50 + sd_quart50),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Quartile 50% mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_quart50_mg

# save into figures folder
ggsave("figures/var_quart_50_mean_MG.png",
       plot = p_quart50_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of quart_75_mean for each MG squirrel
quart75_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_quart75 = mean(quart_75_mean, na.rm = TRUE),
    sd_quart75 = sd(quart_75_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_quart75_mg <- quart75_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_quart75), y = mean_quart75)) +
  geom_pointrange(
    aes(ymin = mean_quart75 - sd_quart75,
        ymax = mean_quart75 + sd_quart75),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Quartile 75% mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_quart75_mg

# save into figures folder
ggsave("figures/var_quart_75_mean_MG.png",
       plot = p_quart75_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of entropy_mean for each MG squirrel
entropy_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_entropy = mean(entropy_mean, na.rm = TRUE),
    sd_entropy = sd(entropy_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_entropy_mg <- entropy_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_entropy), y = mean_entropy)) +
  geom_pointrange(
    aes(ymin = mean_entropy - sd_entropy,
        ymax = mean_entropy + sd_entropy),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Entropy mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_entropy_mg

# save into figures folder
ggsave("figures/var_entropy_mean_MG.png",
       plot = p_entropy_mg,
       width = 8, height = 5, dpi = 300)


# mean + SD of note_rate for each MG squirrel (across calls)
note_rate_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_note_rate = mean(note_rate, na.rm = TRUE),
    sd_note_rate = sd(note_rate, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_note_rate_mg <- note_rate_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_note_rate), y = mean_note_rate)) +
  geom_pointrange(
    aes(ymin = mean_note_rate - sd_note_rate,
        ymax = mean_note_rate + sd_note_rate),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Note rate: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_note_rate_mg

# save it
ggsave("figures/var_note_rate_MG.png",
       plot = p_note_rate_mg,
       width = 8, height = 5, dpi = 300)





```

## Visualization for averages-WM

```{r}

call_avgs_wm <- call_avgs %>%
  left_join(
    sq_data_avgs %>% distinct(sq_ID, site),
    by = "sq_ID"
  ) %>%
  filter(site == "WM")

# mean + SD (across calls) for each WM squirrel
call_duration_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_call_duration = mean(call_duration, na.rm = TRUE),
    sd_call_duration = sd(call_duration, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_call_duration_wm <- call_duration_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_call_duration), y = mean_call_duration)) +
  geom_pointrange(
    aes(ymin = mean_call_duration - sd_call_duration,
        ymax = mean_call_duration + sd_call_duration),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Call duration: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_call_duration_wm

ggsave("figures/var_call_duration_WM.png",
       plot = p_call_duration_wm,
       width = 8, height = 5, dpi = 300)

# note rate
note_rate_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_note_rate = mean(note_rate, na.rm = TRUE),
    sd_note_rate = sd(note_rate, na.rm = TRUE)
  ) %>%
  ungroup()

p_note_rate_wm <- note_rate_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_note_rate), y = mean_note_rate)) +
  geom_pointrange(
    aes(ymin = mean_note_rate - sd_note_rate,
        ymax = mean_note_rate + sd_note_rate),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Note rate: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_note_rate_wm

ggsave("figures/var_note_rate_WM.png",
       plot = p_note_rate_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of duration for each WM squirrel
duration_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_duration = mean(duration, na.rm = TRUE),
    sd_duration = sd(duration, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_duration_wm <- duration_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_duration), y = mean_duration)) +
  geom_pointrange(
    aes(ymin = mean_duration - sd_duration,
        ymax = mean_duration + sd_duration),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Duration: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_duration_wm

# save it
ggsave("figures/var_duration_WM.png",
       plot = p_duration_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of break_s for each WM squirrel
break_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_break = mean(break_s, na.rm = TRUE),
    sd_break = sd(break_s, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_break_wm <- break_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_break), y = mean_break)) +
  geom_pointrange(
    aes(ymin = mean_break - sd_break,
        ymax = mean_break + sd_break),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Break_s: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_break_wm

# save it
ggsave("figures/var_break_s_WM.png",
       plot = p_break_wm,
       width = 8, height = 5, dpi = 300)


# mean + SD (across calls) of rms for each WM squirrel
rms_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_rms = mean(rms, na.rm = TRUE),
    sd_rms = sd(rms, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_rms_wm <- rms_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_rms), y = mean_rms)) +
  geom_pointrange(
    aes(ymin = mean_rms - sd_rms,
        ymax = mean_rms + sd_rms),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "RMS: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_rms_wm

# save it
ggsave("figures/var_rms_WM.png",
       plot = p_rms_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of energy for each WM squirrel
energy_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_energy = mean(energy, na.rm = TRUE),
    sd_energy = sd(energy, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_energy_wm <- energy_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_energy), y = mean_energy)) +
  geom_pointrange(
    aes(ymin = mean_energy - sd_energy,
        ymax = mean_energy + sd_energy),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Energy: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_energy_wm

# save it
ggsave("figures/var_energy_WM.png",
       plot = p_energy_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of peaktopeak for each WM squirrel
peaktopeak_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_peaktopeak = mean(peaktopeak, na.rm = TRUE),
    sd_peaktopeak = sd(peaktopeak, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_peaktopeak_wm <- peaktopeak_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_peaktopeak), y = mean_peaktopeak)) +
  geom_pointrange(
    aes(ymin = mean_peaktopeak - sd_peaktopeak,
        ymax = mean_peaktopeak + sd_peaktopeak),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Peak-to-peak: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_peaktopeak_wm

# save it
ggsave("figures/var_peaktopeak_WM.png",
       plot = p_peaktopeak_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of peak_freq_mean for each WM squirrel
peak_freq_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_peak_freq = mean(peak_freq_mean, na.rm = TRUE),
    sd_peak_freq = sd(peak_freq_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_peak_freq_wm <- peak_freq_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_peak_freq), y = mean_peak_freq)) +
  geom_pointrange(
    aes(ymin = mean_peak_freq - sd_peak_freq,
        ymax = mean_peak_freq + sd_peak_freq),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Peak frequency mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_peak_freq_wm

# save it
ggsave("figures/var_peak_freq_mean_WM.png",
       plot = p_peak_freq_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of peak_ampl_mean for each WM squirrel
peak_ampl_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_peak_ampl = mean(peak_ampl_mean, na.rm = TRUE),
    sd_peak_ampl = sd(peak_ampl_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_peak_ampl_wm <- peak_ampl_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_peak_ampl), y = mean_peak_ampl)) +
  geom_pointrange(
    aes(ymin = mean_peak_ampl - sd_peak_ampl,
        ymax = mean_peak_ampl + sd_peak_ampl),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Peak amplitude mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_peak_ampl_wm

# save it
ggsave("figures/var_peak_ampl_mean_WM.png",
       plot = p_peak_ampl_wm,
       width = 8, height = 5, dpi = 300)


# mean + SD (across calls) of min_freq_mean for each WM squirrel
min_freq_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_min_freq = mean(min_freq_mean, na.rm = TRUE),
    sd_min_freq = sd(min_freq_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_min_freq_wm <- min_freq_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_min_freq), y = mean_min_freq)) +
  geom_pointrange(
    aes(ymin = mean_min_freq - sd_min_freq,
        ymax = mean_min_freq + sd_min_freq),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Min frequency mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_min_freq_wm

# save it
ggsave("figures/var_min_freq_mean_WM.png",
       plot = p_min_freq_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of max_freq_mean for each WM squirrel
max_freq_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_max_freq = mean(max_freq_mean, na.rm = TRUE),
    sd_max_freq = sd(max_freq_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_max_freq_wm <- max_freq_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_max_freq), y = mean_max_freq)) +
  geom_pointrange(
    aes(ymin = mean_max_freq - sd_max_freq,
        ymax = mean_max_freq + sd_max_freq),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Max frequency mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_max_freq_wm

# save it
ggsave("figures/var_max_freq_mean_WM.png",
       plot = p_max_freq_wm,
       width = 8, height = 5, dpi = 300)


# mean + SD (across calls) of bandw_mean for each WM squirrel
bandw_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_bandw = mean(bandw_mean, na.rm = TRUE),
    sd_bandw = sd(bandw_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_bandw_wm <- bandw_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_bandw), y = mean_bandw)) +
  geom_pointrange(
    aes(ymin = mean_bandw - sd_bandw,
        ymax = mean_bandw + sd_bandw),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Bandwidth mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_bandw_wm

# save it
ggsave("figures/var_bandw_mean_WM.png",
       plot = p_bandw_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of quart_25_mean for each WM squirrel
quart25_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_quart25 = mean(quart_25_mean, na.rm = TRUE),
    sd_quart25 = sd(quart_25_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_quart25_wm <- quart25_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_quart25), y = mean_quart25)) +
  geom_pointrange(
    aes(ymin = mean_quart25 - sd_quart25,
        ymax = mean_quart25 + sd_quart25),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Quartile 25% mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_quart25_wm

# save it
ggsave("figures/var_quart_25_mean_WM.png",
       plot = p_quart25_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of quart_50_mean for each WM squirrel
quart50_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_quart50 = mean(quart_50_mean, na.rm = TRUE),
    sd_quart50 = sd(quart_50_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_quart50_wm <- quart50_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_quart50), y = mean_quart50)) +
  geom_pointrange(
    aes(ymin = mean_quart50 - sd_quart50,
        ymax = mean_quart50 + sd_quart50),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Quartile 50% mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_quart50_wm

# save it
ggsave("figures/var_quart_50_mean_WM.png",
       plot = p_quart50_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of quart_75_mean for each WM squirrel
quart75_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_quart75 = mean(quart_75_mean, na.rm = TRUE),
    sd_quart75 = sd(quart_75_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_quart75_wm <- quart75_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_quart75), y = mean_quart75)) +
  geom_pointrange(
    aes(ymin = mean_quart75 - sd_quart75,
        ymax = mean_quart75 + sd_quart75),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Quartile 75% mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_quart75_wm

# save it
ggsave("figures/var_quart_75_mean_WM.png",
       plot = p_quart75_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of entropy_mean for each WM squirrel
entropy_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_entropy = mean(entropy_mean, na.rm = TRUE),
    sd_entropy = sd(entropy_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_entropy_wm <- entropy_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_entropy), y = mean_entropy)) +
  geom_pointrange(
    aes(ymin = mean_entropy - sd_entropy,
        ymax = mean_entropy + sd_entropy),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Entropy mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_entropy_wm

# save it
ggsave("figures/var_entropy_mean_WM.png",
       plot = p_entropy_wm,
       width = 8, height = 5, dpi = 300)


```

## Stats mixed model MG

Do mt graham red squirrels have individual distinctiveness in their calls? Do White Mountain squirrels have distiveness in their calls? I will run a one way ANOVA and then control for inflated p-values by running a FDR/BH sensitivity test.

I will also run a mixed model to better explain variation.

```{r}
########### Run a mixed model ###############

# Mount Graham call-level data for mixed models
mg_df <- call_avgs_mg

library(lme4)

# Just doing one parameter to test 'entropy_mean'
mg_entropy_mod <- lmer(entropy_mean ~ 1 + (1 | sq_ID), data = mg_df)
summary(mg_entropy_mod)

# compute ICC

mg_entropy_vc <- VarCorr(mg_entropy_mod)

mg_var_sq_ID <- as.numeric(mg_entropy_vc$sq_ID[1])
mg_var_resid  <- attr(mg_entropy_vc, "sc")^2

mg_icc_entropy <- mg_var_sq_ID / (mg_var_sq_ID + mg_var_resid)
mg_icc_entropy

# grand means

mg_entropy_effects <- ranef(mg_entropy_mod)$sq_ID[, 1]
mg_entropy_grand_mean <- fixef(mg_entropy_mod)[1]

mg_entropy_means <- mg_entropy_grand_mean + mg_entropy_effects
mg_entropy_means


# Everything looks good, so I will run all parameters

mg_params <- c(
  "call_duration",
  "no_notes",
  "note_rate",
  "duration",
  "break_s",
  "disttomax",
  "rms",
  "energy",
  "peaktopeak",
  "peak_freq_mean",
  "peak_ampl_mean",
  "min_freq_mean",
  "max_freq_mean",
  "bandw_mean",
  "quart_25_mean",
  "quart_50_mean",
  "quart_75_mean",
  "entropy_mean"
)
mg_model_results <- lapply(mg_params, function(param_name) {

  # fit model for one parameter
  mg_mod <- lmer(
    as.formula(paste0(param_name, " ~ 1 + (1 | sq_ID)")),
    data = mg_df
  )

  # variance components -> ICC
  vc <- VarCorr(mg_mod)
  var_sq_ID <- as.numeric(vc$sq_ID[1])
  var_resid  <- attr(vc, "sc")^2
  icc_val <- var_sq_ID / (var_sq_ID + var_resid)

  # return a small list
  list(
    parameter = param_name,
    model = mg_mod,
    var_sq_ID = var_sq_ID,
    var_resid = var_resid,
    icc = icc_val
  )
})
mg_icc_results <- tibble(
  parameter  = sapply(mg_model_results, `[[`, "parameter"),
  var_sq_ID = sapply(mg_model_results, `[[`, "var_sq_ID"),
  var_resid  = sapply(mg_model_results, `[[`, "var_resid"),
  icc        = sapply(mg_model_results, `[[`, "icc")
) %>%
  arrange(desc(icc))

mg_icc_results

write.csv(mg_icc_results, "tables/mg_icc_results.csv", row.names = FALSE)

# plot results
mg_icc_results %>%
  arrange(icc) %>%
  ggplot(aes(x = icc, y = reorder(parameter, icc))) +
  geom_point(na.rm = TRUE) +
  labs(
    x = "ICC (proportion of variance among squirrels)",
    y = "Parameter",
    title = "Mount Graham: individual distinctiveness by parameter (ICC)"
  )


p_mg_icc <- mg_icc_results %>%
  arrange(icc) %>%
  ggplot(aes(x = icc, y = reorder(parameter, icc))) +
  geom_point(na.rm = TRUE) +
  labs(
    x = "ICC (proportion of variance among squirrels)",
    y = "Parameter",
    title = "Mount Graham: ICC for each acoustic parameter"
  )

p_mg_icc
ggsave("figures/mg_icc_dotplot.png", plot = p_mg_icc, width = 8, height = 5, dpi = 300)

```

## Stats mixed model WM

Run a mixed model for the white mountain squirrels

```{r}
# White Mountains: mixed models + ICC table + dotplot

#  WM call-level data
wm_df <- call_avgs_wm

#  Same parameter list
wm_params <- c(
  "call_duration",
  "no_notes",
  "note_rate",
  "duration",
  "break_s",
  "disttomax",
  "rms",
  "energy",
  "peaktopeak",
  "peak_freq_mean",
  "peak_ampl_mean",
  "min_freq_mean",
  "max_freq_mean",
  "bandw_mean",
  "quart_25_mean",
  "quart_50_mean",
  "quart_75_mean",
  "entropy_mean"
)

#  Run all WM models + ICC
wm_model_results <- lapply(wm_params, function(param_name) {

  wm_mod <- lmer(
    as.formula(paste0(param_name, " ~ 1 + (1 | sq_ID)")),
    data = wm_df
  )

  vc <- VarCorr(wm_mod)
  var_sq_ID <- as.numeric(vc$sq_ID[1])
  var_resid  <- attr(vc, "sc")^2
  icc_val <- var_sq_ID / (var_sq_ID + var_resid)

  list(
    parameter = param_name,
    model = wm_mod,
    var_sq_ID = var_sq_ID,
    var_resid = var_resid,
    icc = icc_val
  )
})

#  WM ICC results table
wm_icc_results <- tibble(
  parameter  = sapply(wm_model_results, `[[`, "parameter"),
  var_sq_ID = sapply(wm_model_results, `[[`, "var_sq_ID"),
  var_resid  = sapply(wm_model_results, `[[`, "var_resid"),
  icc        = sapply(wm_model_results, `[[`, "icc")
) %>%
  arrange(desc(icc))

wm_icc_results
write.csv(wm_icc_results, "tables/wm_icc_results.csv", row.names = FALSE)

# WM ICC dot plot
p_wm_icc <- wm_icc_results %>%
  arrange(icc) %>%
  ggplot(aes(x = icc, y = reorder(parameter, icc))) +
  geom_point(na.rm = TRUE) +
  labs(
    x = "ICC (proportion of variance among squirrels)",
    y = "Parameter",
    title = "White Mountains: individual distinctiveness by parameter (ICC)"
  )

p_wm_icc
ggsave("tables/wm_icc_dotplot.png", plot = p_wm_icc, width = 8, height = 5, dpi = 300)

```

## Visualization

```{r}

```

## Stats 

```{r}

```

## Population differences

Split MG from WM
