---
title: "Identity information analysis"
author: "Nikki Reck"
format: html
editor: visual
---

## Background

Analysis of acoustic identity information of Mount Graham red squirrels and White Mountain red squirrels.

## Load Data

Load in packages and the data.

```{r}
library(lme4)
library(effectsize)
library(tidyverse)

#### Load Data Avisoft rattle detections spreadsheet ###

detections <- read_csv("data_raw/detections10.csv") %>%
  rename(
    note_number    = `#`,
    break_s        = `break`,
    start_time     = `start time`,
    end_time       = `end time`,

    peak_freq_mean = `peak freq(mean)`,
    peak_ampl_mean = `peak ampl(mean)`,
    min_freq_mean  = `min freq(mean)`,
    max_freq_mean  = `max freq(mean)`,
    bandw_mean     = `bandw(mean)`,

    quart_25_mean  = `quart 25(mean)`,
    quart_50_mean  = `quart 50(mean)`,
    quart_75_mean  = `quart 75(mean)`,
    entropy_mean   = `entropy(mean)`
  )

#### Load squirrel data spreadsheet ####
combined <- read_csv("data_raw/combined_sq_data_new.csv")




# Clean break_s column 
detections <- detections %>%
  mutate(
    # Clean break_s: "_" = missing, then make numeric
    break_s = as.character(break_s),
    break_s = na_if(break_s, "_"),
    break_s = as.numeric(break_s),
   
    # Create uniqueid from filename
    uniqueid = filename %>%
      str_replace("^.*[\\\\/]", "") %>%   # drop any path (works for \ and /)
      str_remove("\\.wav$") %>%          # drop extension
      if_else(str_count(., "_") > 2, str_replace(., "^[^_]*_", ""), .)  # drop leading prefix if present
  )

# make sure all dates, TOD, are consistent
combined <- combined %>%
  mutate(
    # TOD as 4-digit text (works whether TOD is numeric or already "0645")
    TOD = str_pad(as.character(TOD), width = 4, pad = "0"),

    # parse date no matter whether it's "06/20/25", "06/20/2025", or "06202025"
    date_parsed = case_when(
      str_detect(as.character(date), "^\\d{2}/\\d{2}/\\d{2}$") ~ as.Date(as.character(date), format = "%m/%d/%y"),
      str_detect(as.character(date), "^\\d{2}/\\d{2}/\\d{4}$") ~ as.Date(as.character(date), format = "%m/%d/%Y"),
      str_detect(as.character(date), "^\\d{8}$")               ~ as.Date(as.character(date), format = "%m%d%Y"),
      TRUE                                                     ~ NA_Date_
    ),

    uniqueid = paste0(midden, "_", format(date_parsed, "%m%d%Y"), "_", TOD)
  )

# Merge Data (Keep all recordings)
sq_rattle_data <- merge(detections, combined, by.x = "uniqueid", by.y = "uniqueid", all = TRUE)

# Calculate note rate 
sq_rattle_data <- sq_rattle_data %>%
  group_by(uniqueid, rattle_no) %>% 
  mutate(
    note_rate = max(note_number, na.rm = TRUE) /
                as.numeric(first(call_duration))
  ) %>%
  ungroup()

```

```{r}

```

## Distance bias check on parameters

```{r}
# Create a log_distance variable to model attenuation correctly
sq_rattle_data <- sq_rattle_data %>%
  mutate(log_distance = log(distance))

# excluding note_number
params <- c(
  "call_duration",
  "note_rate",
  "duration",
  "break_s",
  "disttomax",
  "rms",
  "energy",
  "peaktopeak",
  "peak_freq_mean",
  "peak_ampl_mean",
  "min_freq_mean",
  "max_freq_mean",
  "bandw_mean",
  "quart_25_mean",
  "quart_50_mean",
  "quart_75_mean",
  "entropy_mean"
)

# --- Visual Check for Biological Distance Bias ---

for (p in params) {
  dist_bias_plot <- ggplot(
    sq_rattle_data,
    aes(x = log_distance, y = !!sym(p), color = site)
  ) +
    geom_point(alpha = 0.5, na.rm = TRUE) +
    geom_smooth(method = "lm", se = TRUE, na.rm = TRUE, color = "black") + # Overall trend line
    labs(
      title = paste("Visual Distance Bias Check:", p),
      x = "Log Distance (m)",
      y = p,
      color = "Population"
    ) +
    theme_bw() +
    theme(legend.position = "top")

  print(dist_bias_plot)
 
  # Ensure the figures/distance_bias folder exists, then save the plot
  if(!dir.exists("figures/distance_bias")) {
    dir.create("figures/distance_bias", recursive = TRUE)
  }
  ggsave(
    filename = paste0("figures/distance_bias/dist_bias_", p, ".png"),
    plot = dist_bias_plot,
    width = 7,
    height = 5,
    dpi = 300
  )
}

# Dist effects for rms, peaktopeak, peak amp mean
```

## Call level Averages

Taking the averages of each of the calls per animal. Fix any oddities.

```{r}

# Call-level averages by sq_ID (one row per call)

call_avgs <- sq_rattle_data %>%
  group_by(sq_ID, site, filename) %>%
  summarise(
    no_notes = n(),
    call_duration = first(call_duration),
    distance = first(distance),
    log_distance = first(log_distance),
    note_rate = no_notes / call_duration,

    duration = mean(duration, na.rm = TRUE),
    break_s = mean(break_s, na.rm = TRUE),
    disttomax = mean(disttomax, na.rm = TRUE),

    rms = mean(rms, na.rm = TRUE),
    energy = mean(energy, na.rm = TRUE),
    peaktopeak = mean(peaktopeak, na.rm = TRUE),

    peak_freq_mean = mean(peak_freq_mean, na.rm = TRUE),
    peak_ampl_mean = mean(peak_ampl_mean, na.rm = TRUE),
    min_freq_mean = mean(min_freq_mean, na.rm = TRUE),
    max_freq_mean = mean(max_freq_mean, na.rm = TRUE),
    bandw_mean = mean(bandw_mean, na.rm = TRUE),

    quart_25_mean = mean(quart_25_mean, na.rm = TRUE),
    quart_50_mean = mean(quart_50_mean, na.rm = TRUE),
    quart_75_mean = mean(quart_75_mean, na.rm = TRUE),

    entropy_mean = mean(entropy_mean, na.rm = TRUE)
  ) %>%
  ungroup()
call_avgs

names(call_avgs)



```

## Squirrel level averages

Taking the averages of all the calls per animal.

```{r}


sq_avgs <- call_avgs %>%
  group_by(sq_ID, site) %>%
  summarise(
    n_calls = n(),

    # averages of call-level traits
    call_duration = mean(call_duration, na.rm = TRUE),
    no_notes = mean(no_notes, na.rm = TRUE),
    note_rate = mean(note_rate, na.rm = TRUE),

    duration = mean(duration, na.rm = TRUE),
    break_s = mean(break_s, na.rm = TRUE),
    disttomax = mean(disttomax, na.rm = TRUE),

    rms = mean(rms, na.rm = TRUE),
    energy = mean(energy, na.rm = TRUE),
    peaktopeak = mean(peaktopeak, na.rm = TRUE),

    peak_freq_mean = mean(peak_freq_mean, na.rm = TRUE),
    peak_ampl_mean = mean(peak_ampl_mean, na.rm = TRUE),
    min_freq_mean = mean(min_freq_mean, na.rm = TRUE),
    max_freq_mean = mean(max_freq_mean, na.rm = TRUE),
    bandw_mean = mean(bandw_mean, na.rm = TRUE),

    quart_25_mean = mean(quart_25_mean, na.rm = TRUE),
    quart_50_mean = mean(quart_50_mean, na.rm = TRUE),
    quart_75_mean = mean(quart_75_mean, na.rm = TRUE),

    entropy_mean = mean(entropy_mean, na.rm = TRUE)
  ) %>%
  ungroup()
sq_avgs
names(sq_avgs)



```

## 

```{r}


```

## Call avgs MG- visuals

Create figures for a clear look into data.

```{r}

# Just going to look at figs for mt graham squirrels.

# using site, keep only Mount Graham call averages
call_avgs_mg <- call_avgs %>%
  filter(site == "MG")



# for MG squirrels, calculate the mean and SD of call_duration across calls
call_duration_by_sq_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_call_duration = mean(call_duration, na.rm = TRUE),
    sd_call_duration = sd(call_duration, na.rm = TRUE)
  ) %>%
  ungroup()

p_call_duration_mg <- call_duration_by_sq_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_call_duration), y = mean_call_duration)) +
  geom_pointrange(
    aes(ymin = mean_call_duration - sd_call_duration,
        ymax = mean_call_duration + sd_call_duration),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Call duration: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_call_duration_mg

ggsave("figures/var_call_duration_MG.png",
       plot = p_call_duration_mg,
       width = 8, height = 5, dpi = 300)

#  for MG squirrels, calculate mean and SD of note_rate across calls
note_rate_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_note_rate = mean(note_rate, na.rm = TRUE),
    sd_note_rate = sd(note_rate, na.rm = TRUE)
  ) %>%
  ungroup()
p_note_rate_mg <- note_rate_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_note_rate), y = mean_note_rate)) +
  geom_pointrange(
    aes(ymin = mean_note_rate - sd_note_rate,
        ymax = mean_note_rate + sd_note_rate),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Note rate: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_note_rate_mg
ggsave("figures/var_note_rate_MG.png",
       plot = p_note_rate_mg,
       width = 8, height = 5, dpi = 300)

# duration param MG
# mean + SD of duration for each MG squirrel
duration_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_duration = mean(duration, na.rm = TRUE),
    sd_duration = sd(duration, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_duration_mg <- duration_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_duration), y = mean_duration)) +
  geom_pointrange(
    aes(ymin = mean_duration - sd_duration,
        ymax = mean_duration + sd_duration),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Duration: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_duration_mg

# save into figures folder
ggsave("figures/var_duration_MG.png",
       plot = p_duration_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of break_s for each MG squirrel
break_s_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_break_s = mean(break_s, na.rm = TRUE),
    sd_break_s = sd(break_s, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_break_s_mg <- break_s_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_break_s), y = mean_break_s)) +
  geom_pointrange(
    aes(ymin = mean_break_s - sd_break_s,
        ymax = mean_break_s + sd_break_s),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "break_s: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_break_s_mg

# save into figures folder
ggsave("figures/var_break_s_MG.png",
       plot = p_break_s_mg,
       width = 8, height = 5, dpi = 300)


# mean + SD of disttomax for each MG squirrel
disttomax_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_disttomax = mean(disttomax, na.rm = TRUE),
    sd_disttomax = sd(disttomax, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_disttomax_mg <- disttomax_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_disttomax), y = mean_disttomax)) +
  geom_pointrange(
    aes(ymin = mean_disttomax - sd_disttomax,
        ymax = mean_disttomax + sd_disttomax),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "disttomax: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_disttomax_mg

# save into figures folder
ggsave("figures/var_disttomax_MG.png",
       plot = p_disttomax_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of rms for each MG squirrel
rms_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_rms = mean(rms, na.rm = TRUE),
    sd_rms = sd(rms, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_rms_mg <- rms_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_rms), y = mean_rms)) +
  geom_pointrange(
    aes(ymin = mean_rms - sd_rms,
        ymax = mean_rms + sd_rms),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "RMS: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_rms_mg

# save into figures folder
ggsave("figures/var_rms_MG.png",
       plot = p_rms_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of energy for each MG squirrel
energy_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_energy = mean(energy, na.rm = TRUE),
    sd_energy = sd(energy, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_energy_mg <- energy_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_energy), y = mean_energy)) +
  geom_pointrange(
    aes(ymin = mean_energy - sd_energy,
        ymax = mean_energy + sd_energy),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Energy: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_energy_mg

# save into figures folder
ggsave("figures/var_energy_MG.png",
       plot = p_energy_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of peaktopeak for each MG squirrel
peaktopeak_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_peaktopeak = mean(peaktopeak, na.rm = TRUE),
    sd_peaktopeak = sd(peaktopeak, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_peaktopeak_mg <- peaktopeak_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_peaktopeak), y = mean_peaktopeak)) +
  geom_pointrange(
    aes(ymin = mean_peaktopeak - sd_peaktopeak,
        ymax = mean_peaktopeak + sd_peaktopeak),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Peak-to-peak: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_peaktopeak_mg

# save into figures folder
ggsave("figures/var_peaktopeak_MG.png",
       plot = p_peaktopeak_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of peak_freq_mean for each MG squirrel
peak_freq_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_peak_freq = mean(peak_freq_mean, na.rm = TRUE),
    sd_peak_freq = sd(peak_freq_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_peak_freq_mg <- peak_freq_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_peak_freq), y = mean_peak_freq)) +
  geom_pointrange(
    aes(ymin = mean_peak_freq - sd_peak_freq,
        ymax = mean_peak_freq + sd_peak_freq),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Peak frequency mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_peak_freq_mg

# save into figures folder
ggsave("figures/var_peak_freq_mean_MG.png",
       plot = p_peak_freq_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of peak_ampl_mean for each MG squirrel
peak_ampl_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_peak_ampl = mean(peak_ampl_mean, na.rm = TRUE),
    sd_peak_ampl = sd(peak_ampl_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_peak_ampl_mg <- peak_ampl_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_peak_ampl), y = mean_peak_ampl)) +
  geom_pointrange(
    aes(ymin = mean_peak_ampl - sd_peak_ampl,
        ymax = mean_peak_ampl + sd_peak_ampl),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Peak amplitude mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_peak_ampl_mg

# save into figures folder
ggsave("figures/var_peak_ampl_mean_MG.png",
       plot = p_peak_ampl_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of min_freq_mean for each MG squirrel
min_freq_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_min_freq = mean(min_freq_mean, na.rm = TRUE),
    sd_min_freq = sd(min_freq_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_min_freq_mg <- min_freq_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_min_freq), y = mean_min_freq)) +
  geom_pointrange(
    aes(ymin = mean_min_freq - sd_min_freq,
        ymax = mean_min_freq + sd_min_freq),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Min frequency mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_min_freq_mg

# save into figures folder
ggsave("figures/var_min_freq_mean_MG.png",
       plot = p_min_freq_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of max_freq_mean for each MG squirrel
max_freq_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_max_freq = mean(max_freq_mean, na.rm = TRUE),
    sd_max_freq = sd(max_freq_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_max_freq_mg <- max_freq_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_max_freq), y = mean_max_freq)) +
  geom_pointrange(
    aes(ymin = mean_max_freq - sd_max_freq,
        ymax = mean_max_freq + sd_max_freq),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Max frequency mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_max_freq_mg

# save into figures folder
ggsave("figures/var_max_freq_mean_MG.png",
       plot = p_max_freq_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of bandw_mean for each MG squirrel
bandw_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_bandw = mean(bandw_mean, na.rm = TRUE),
    sd_bandw = sd(bandw_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_bandw_mg <- bandw_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_bandw), y = mean_bandw)) +
  geom_pointrange(
    aes(ymin = mean_bandw - sd_bandw,
        ymax = mean_bandw + sd_bandw),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Bandwidth mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_bandw_mg

# save into figures folder
ggsave("figures/var_bandw_mean_MG.png",
       plot = p_bandw_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of quart_25_mean for each MG squirrel
quart25_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_quart25 = mean(quart_25_mean, na.rm = TRUE),
    sd_quart25 = sd(quart_25_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_quart25_mg <- quart25_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_quart25), y = mean_quart25)) +
  geom_pointrange(
    aes(ymin = mean_quart25 - sd_quart25,
        ymax = mean_quart25 + sd_quart25),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Quartile 25% mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_quart25_mg

# save into figures folder
ggsave("figures/var_quart_25_mean_MG.png",
       plot = p_quart25_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of quart_50_mean for each MG squirrel
quart50_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_quart50 = mean(quart_50_mean, na.rm = TRUE),
    sd_quart50 = sd(quart_50_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_quart50_mg <- quart50_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_quart50), y = mean_quart50)) +
  geom_pointrange(
    aes(ymin = mean_quart50 - sd_quart50,
        ymax = mean_quart50 + sd_quart50),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Quartile 50% mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_quart50_mg

# save into figures folder
ggsave("figures/var_quart_50_mean_MG.png",
       plot = p_quart50_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of quart_75_mean for each MG squirrel
quart75_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_quart75 = mean(quart_75_mean, na.rm = TRUE),
    sd_quart75 = sd(quart_75_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_quart75_mg <- quart75_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_quart75), y = mean_quart75)) +
  geom_pointrange(
    aes(ymin = mean_quart75 - sd_quart75,
        ymax = mean_quart75 + sd_quart75),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Quartile 75% mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_quart75_mg

# save into figures folder
ggsave("figures/var_quart_75_mean_MG.png",
       plot = p_quart75_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of entropy_mean for each MG squirrel
entropy_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_entropy = mean(entropy_mean, na.rm = TRUE),
    sd_entropy = sd(entropy_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_entropy_mg <- entropy_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_entropy), y = mean_entropy)) +
  geom_pointrange(
    aes(ymin = mean_entropy - sd_entropy,
        ymax = mean_entropy + sd_entropy),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Entropy mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_entropy_mg

# save into figures folder
ggsave("figures/var_entropy_mean_MG.png",
       plot = p_entropy_mg,
       width = 8, height = 5, dpi = 300)


# mean + SD of note_rate for each MG squirrel (across calls)
note_rate_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_note_rate = mean(note_rate, na.rm = TRUE),
    sd_note_rate = sd(note_rate, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_note_rate_mg <- note_rate_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_note_rate), y = mean_note_rate)) +
  geom_pointrange(
    aes(ymin = mean_note_rate - sd_note_rate,
        ymax = mean_note_rate + sd_note_rate),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Note rate: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_note_rate_mg

# save it
ggsave("figures/var_note_rate_MG.png",
       plot = p_note_rate_mg,
       width = 8, height = 5, dpi = 300)





```

## Call avgs WM- visuals

```{r}

call_avgs_wm <- call_avgs %>%
  filter(site == "WM")


# mean + SD (across calls) for each WM squirrel
call_duration_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_call_duration = mean(call_duration, na.rm = TRUE),
    sd_call_duration = sd(call_duration, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_call_duration_wm <- call_duration_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_call_duration), y = mean_call_duration)) +
  geom_pointrange(
    aes(ymin = mean_call_duration - sd_call_duration,
        ymax = mean_call_duration + sd_call_duration),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Call duration: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_call_duration_wm

ggsave("figures/var_call_duration_WM.png",
       plot = p_call_duration_wm,
       width = 8, height = 5, dpi = 300)

# note rate
note_rate_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_note_rate = mean(note_rate, na.rm = TRUE),
    sd_note_rate = sd(note_rate, na.rm = TRUE)
  ) %>%
  ungroup()

p_note_rate_wm <- note_rate_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_note_rate), y = mean_note_rate)) +
  geom_pointrange(
    aes(ymin = mean_note_rate - sd_note_rate,
        ymax = mean_note_rate + sd_note_rate),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Note rate: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_note_rate_wm

ggsave("figures/var_note_rate_WM.png",
       plot = p_note_rate_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of duration for each WM squirrel
duration_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_duration = mean(duration, na.rm = TRUE),
    sd_duration = sd(duration, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_duration_wm <- duration_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_duration), y = mean_duration)) +
  geom_pointrange(
    aes(ymin = mean_duration - sd_duration,
        ymax = mean_duration + sd_duration),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Duration: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_duration_wm

# save it
ggsave("figures/var_duration_WM.png",
       plot = p_duration_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of break_s for each WM squirrel
break_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_break = mean(break_s, na.rm = TRUE),
    sd_break = sd(break_s, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_break_wm <- break_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_break), y = mean_break)) +
  geom_pointrange(
    aes(ymin = mean_break - sd_break,
        ymax = mean_break + sd_break),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Break_s: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_break_wm

# save it
ggsave("figures/var_break_s_WM.png",
       plot = p_break_wm,
       width = 8, height = 5, dpi = 300)


# mean + SD (across calls) of rms for each WM squirrel
rms_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_rms = mean(rms, na.rm = TRUE),
    sd_rms = sd(rms, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_rms_wm <- rms_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_rms), y = mean_rms)) +
  geom_pointrange(
    aes(ymin = mean_rms - sd_rms,
        ymax = mean_rms + sd_rms),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "RMS: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_rms_wm

# save it
ggsave("figures/var_rms_WM.png",
       plot = p_rms_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of energy for each WM squirrel
energy_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_energy = mean(energy, na.rm = TRUE),
    sd_energy = sd(energy, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_energy_wm <- energy_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_energy), y = mean_energy)) +
  geom_pointrange(
    aes(ymin = mean_energy - sd_energy,
        ymax = mean_energy + sd_energy),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Energy: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_energy_wm

# save it
ggsave("figures/var_energy_WM.png",
       plot = p_energy_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of peaktopeak for each WM squirrel
peaktopeak_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_peaktopeak = mean(peaktopeak, na.rm = TRUE),
    sd_peaktopeak = sd(peaktopeak, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_peaktopeak_wm <- peaktopeak_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_peaktopeak), y = mean_peaktopeak)) +
  geom_pointrange(
    aes(ymin = mean_peaktopeak - sd_peaktopeak,
        ymax = mean_peaktopeak + sd_peaktopeak),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Peak-to-peak: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_peaktopeak_wm

# save it
ggsave("figures/var_peaktopeak_WM.png",
       plot = p_peaktopeak_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of peak_freq_mean for each WM squirrel
peak_freq_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_peak_freq = mean(peak_freq_mean, na.rm = TRUE),
    sd_peak_freq = sd(peak_freq_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_peak_freq_wm <- peak_freq_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_peak_freq), y = mean_peak_freq)) +
  geom_pointrange(
    aes(ymin = mean_peak_freq - sd_peak_freq,
        ymax = mean_peak_freq + sd_peak_freq),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Peak frequency mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_peak_freq_wm

# save it
ggsave("figures/var_peak_freq_mean_WM.png",
       plot = p_peak_freq_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of peak_ampl_mean for each WM squirrel
peak_ampl_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_peak_ampl = mean(peak_ampl_mean, na.rm = TRUE),
    sd_peak_ampl = sd(peak_ampl_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_peak_ampl_wm <- peak_ampl_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_peak_ampl), y = mean_peak_ampl)) +
  geom_pointrange(
    aes(ymin = mean_peak_ampl - sd_peak_ampl,
        ymax = mean_peak_ampl + sd_peak_ampl),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Peak amplitude mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_peak_ampl_wm

# save it
ggsave("figures/var_peak_ampl_mean_WM.png",
       plot = p_peak_ampl_wm,
       width = 8, height = 5, dpi = 300)


# mean + SD (across calls) of min_freq_mean for each WM squirrel
min_freq_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_min_freq = mean(min_freq_mean, na.rm = TRUE),
    sd_min_freq = sd(min_freq_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_min_freq_wm <- min_freq_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_min_freq), y = mean_min_freq)) +
  geom_pointrange(
    aes(ymin = mean_min_freq - sd_min_freq,
        ymax = mean_min_freq + sd_min_freq),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Min frequency mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_min_freq_wm

# save it
ggsave("figures/var_min_freq_mean_WM.png",
       plot = p_min_freq_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of max_freq_mean for each WM squirrel
max_freq_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_max_freq = mean(max_freq_mean, na.rm = TRUE),
    sd_max_freq = sd(max_freq_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_max_freq_wm <- max_freq_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_max_freq), y = mean_max_freq)) +
  geom_pointrange(
    aes(ymin = mean_max_freq - sd_max_freq,
        ymax = mean_max_freq + sd_max_freq),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Max frequency mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_max_freq_wm

# save it
ggsave("figures/var_max_freq_mean_WM.png",
       plot = p_max_freq_wm,
       width = 8, height = 5, dpi = 300)


# mean + SD (across calls) of bandw_mean for each WM squirrel
bandw_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_bandw = mean(bandw_mean, na.rm = TRUE),
    sd_bandw = sd(bandw_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_bandw_wm <- bandw_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_bandw), y = mean_bandw)) +
  geom_pointrange(
    aes(ymin = mean_bandw - sd_bandw,
        ymax = mean_bandw + sd_bandw),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Bandwidth mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_bandw_wm

# save it
ggsave("figures/var_bandw_mean_WM.png",
       plot = p_bandw_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of quart_25_mean for each WM squirrel
quart25_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_quart25 = mean(quart_25_mean, na.rm = TRUE),
    sd_quart25 = sd(quart_25_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_quart25_wm <- quart25_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_quart25), y = mean_quart25)) +
  geom_pointrange(
    aes(ymin = mean_quart25 - sd_quart25,
        ymax = mean_quart25 + sd_quart25),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Quartile 25% mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_quart25_wm

# save it
ggsave("figures/var_quart_25_mean_WM.png",
       plot = p_quart25_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of quart_50_mean for each WM squirrel
quart50_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_quart50 = mean(quart_50_mean, na.rm = TRUE),
    sd_quart50 = sd(quart_50_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_quart50_wm <- quart50_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_quart50), y = mean_quart50)) +
  geom_pointrange(
    aes(ymin = mean_quart50 - sd_quart50,
        ymax = mean_quart50 + sd_quart50),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Quartile 50% mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_quart50_wm

# save it
ggsave("figures/var_quart_50_mean_WM.png",
       plot = p_quart50_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of quart_75_mean for each WM squirrel
quart75_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_quart75 = mean(quart_75_mean, na.rm = TRUE),
    sd_quart75 = sd(quart_75_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_quart75_wm <- quart75_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_quart75), y = mean_quart75)) +
  geom_pointrange(
    aes(ymin = mean_quart75 - sd_quart75,
        ymax = mean_quart75 + sd_quart75),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Quartile 75% mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_quart75_wm

# save it
ggsave("figures/var_quart_75_mean_WM.png",
       plot = p_quart75_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of entropy_mean for each WM squirrel
entropy_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_entropy = mean(entropy_mean, na.rm = TRUE),
    sd_entropy = sd(entropy_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_entropy_wm <- entropy_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_entropy), y = mean_entropy)) +
  geom_pointrange(
    aes(ymin = mean_entropy - sd_entropy,
        ymax = mean_entropy + sd_entropy),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Entropy mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_entropy_wm

# save it
ggsave("figures/var_entropy_mean_WM.png",
       plot = p_entropy_wm,
       width = 8, height = 5, dpi = 300)


```

## Distance Cleaning

```{r}
# Creating a vector that contains all the parameters that show distance sensitivity. All my other parameters will run normally with distance correction. (params)

dist_sensitive_params <- c(
  "call_duration",
  "rms",
  "peak_ampl_mean",
  "peaktopeak",
  "quart_75_mean"
)

# create a distance correction function

apply_distance_correction <- function(df, params_to_correct) {
  df_cleaned <- as.data.frame(df) # Force it to a plain dataframe to fix names() bugs
 
  for (p in params_to_correct) {
    if (p %in% colnames(df_cleaned)) {
      # 1. Fit linear model (parameter ~ log_distance)
      formula_str <- paste0(p, " ~ log_distance")
      fit <- lm(as.formula(formula_str), data = df_cleaned, na.action = na.exclude)
     
      # 2. Get the predicted value at the average distance
      mean_dist <- mean(df_cleaned$log_distance, na.rm = TRUE)
      pred_at_mean <- predict(fit, newdata = data.frame(log_distance = mean_dist))
     
      # 3. Extract individual residuals (the "fingerprint")
      resids <- resid(fit)
     
      # 4. Corrected value = prediction at average + residuals
      df_cleaned[[p]] <- pred_at_mean + resids
    }
  }
  return(df_cleaned)
}

# Clean both datasets
call_avgs_mg_cleaned <- apply_distance_correction(call_avgs_mg, dist_sensitive_params)
call_avgs_wm_cleaned <- apply_distance_correction(call_avgs_wm, dist_sensitive_params)

```

## ICC MG

```{r}
########### Run a mixed model ###############

# Mount Graham call-level data for mixed models (distance adjusted)
mg_df <- call_avgs_mg_cleaned

mg_model_results <- lapply(params, function(param_name) {

  mod_formula <- paste0(param_name, " ~ 1 + (1 | sq_ID)")

  # fit model for one parameter
  mg_mod <- lmer(
    as.formula(mod_formula),
    data = mg_df
  )

  # variance components -> ICC
  vc <- VarCorr(mg_mod)
  var_sq_ID <- as.numeric(vc$sq_ID[1])
  var_resid  <- attr(vc, "sc")^2
  icc_val <- var_sq_ID / (var_sq_ID + var_resid)

  # return a small list
  list(
    parameter = param_name,
    model = mg_mod,
    var_sq_ID = var_sq_ID,
    var_resid = var_resid,
    icc = icc_val
  )
})
mg_icc_results <- tibble(
  parameter  = sapply(mg_model_results, `[[`, "parameter"),
  var_sq_ID = sapply(mg_model_results, `[[`, "var_sq_ID"),
  var_resid  = sapply(mg_model_results, `[[`, "var_resid"),
  icc        = sapply(mg_model_results, `[[`, "icc")
) %>%
  arrange(desc(icc))

mg_icc_results

write.csv(mg_icc_results, "tables/mg_icc_results.csv", row.names = FALSE)

# plot results
mg_icc_results %>%
  arrange(icc) %>%
  ggplot(aes(x = icc, y = reorder(parameter, icc))) +
  geom_point(na.rm = TRUE) +
  labs(
    x = "ICC (proportion of variance among squirrels)",
    y = "Parameter",
    title = "Mount Graham: individual distinctiveness by parameter (ICC)"
  )


p_mg_icc <- mg_icc_results %>%
  arrange(icc) %>%
  ggplot(aes(x = icc, y = reorder(parameter, icc))) +
  geom_point(na.rm = TRUE) +
  labs(
    x = "ICC (proportion of variance among squirrels)",
    y = "Parameter",
    title = "Mount Graham: ICC for each acoustic parameter"
  )

p_mg_icc
ggsave("figures/mg_icc_dotplot.png", plot = p_mg_icc, width = 8, height = 5, dpi = 300)

```

## ICC WM

```{r}
# White Mountains: mixed models (distance adjusted)

#  WM call-level data (CLEANED)
wm_df <- call_avgs_wm_cleaned

#  Run all WM models + ICC
wm_model_results <- lapply(params, function(param_name) {

  wm_mod_formula <- paste0(param_name, " ~ 1 + (1 | sq_ID)")

  # fit model for one parameter
  wm_mod <- lmer(
    as.formula(wm_mod_formula),
    data = wm_df
  )

  vc <- VarCorr(wm_mod)
  var_sq_ID <- as.numeric(vc$sq_ID[1])
  var_resid  <- attr(vc, "sc")^2
  icc_val <- var_sq_ID / (var_sq_ID + var_resid)

  list(
    parameter = param_name,
    model = wm_mod,
    var_sq_ID = var_sq_ID,
    var_resid = var_resid,
    icc = icc_val
  )
})

#  WM ICC results table
wm_icc_results <- tibble(
  parameter  = sapply(wm_model_results, `[[`, "parameter"),
  var_sq_ID = sapply(wm_model_results, `[[`, "var_sq_ID"),
  var_resid  = sapply(wm_model_results, `[[`, "var_resid"),
  icc        = sapply(wm_model_results, `[[`, "icc")
) %>%
  arrange(desc(icc))

wm_icc_results
write.csv(wm_icc_results, "tables/wm_icc_results.csv", row.names = FALSE)

# WM ICC dot plot
p_wm_icc <- wm_icc_results %>%
  arrange(icc) %>%
  ggplot(aes(x = icc, y = reorder(parameter, icc))) +
  geom_point(na.rm = TRUE) +
  labs(
    x = "ICC (proportion of variance among squirrels)",
    y = "Parameter",
    title = "White Mountains: individual distinctiveness by parameter (ICC)"
  )

p_wm_icc
ggsave("figures/wm_icc_dotplot.png", plot = p_wm_icc, width = 8, height = 5, dpi = 300)

```

## ICC bootstrapping

```{r}
# ------------------------------------------------------------------------------
# ICC Analysis with Bootstrapping
# ------------------------------------------------------------------------------

# EXPLANATION OF BOOTSTRAPPING:
# Bootstrapping is a resampling method used to estimate statistics (like ICC)
# and their uncertainty (Confidence Intervals) by randomly sampling the observed data
# with replacement. It simulates repeating the experiment many times to see
# how much the result varies. This is especially useful for biological data
# where distributions may not be perfectly normal.

# Function to calculate ICC from a linear mixed model
# ICC = variance_among_groups / (variance_among_groups + variance_within_groups)

calc_icc_lmer <- function(data, indices, formula) {
  d <- data[indices, ] # resample data
 
  # Fit the model with the resampled data
  # use tryCatch to handle cases where the model might not converge on a random sample
  fit <- tryCatch({
    lme4::lmer(formula, data = d, REML = TRUE) # REML is simpler/faster for this
  }, error = function(e) return(NULL))
 
  if (is.null(fit)) return(NA)
 
  # Extract variance components
  vc <- as.data.frame(lme4::VarCorr(fit))
  var_among  <- vc$vcov[vc$grp == "sq_ID"]
  var_within <- vc$vcov[vc$grp == "Residual"]
 
  # Calculate ICC
  icc <- var_among / (var_among + var_within)
  return(icc)
}

# Wrapper function to run bootstrap for a specific parameter and site
run_bootstrap_icc <- function(df, param, n_reps = 1000) {
  # Formula: parameter depends on random effect of sq_ID
  # distance corrected with log_distance
  f <- as.formula(paste(param, "~ log_distance + (1 | sq_ID)"))
 
  boot_res <- boot::boot(data = df, statistic = calc_icc_lmer, R = n_reps, formula = f)
 
  # Get Confidence Interval (Basic or Percentile)
  ci <- tryCatch({
    boot::boot.ci(boot_res, type = "perc")$percent[4:5]
  }, error = function(e) c(NA, NA))
 
  return(list(
    icc = boot_res$t0,       # Observed ICC
    ci_lower = ci[1],        # Lower CI
    ci_upper = ci[2]         # Upper CI
  ))
}

# Run ICC analysis for all parameters for both sites
icc_results <- data.frame()

for (site_name in unique(sq_avgs$site)) {
  if (is.na(site_name)) next # skip if NA site
 
  site_data <- call_avgs %>% filter(site == site_name)
 
  for (p in params) {
    if(p == "no_notes" || p == "call_duration") next # Skip these if not acoustic parameters
   
    cat("Bootstrapping ICC for:", site_name, "-", p, "\n")
   
    res <- run_bootstrap_icc(site_data, p, n_reps = 500) # 500 reps for speed, use 1000+ for final
   
    icc_results <- rbind(icc_results, data.frame(
      site = site_name,
      parameter = p,
      ICC = res$icc,
      CI_Lower = res$ci_lower,
      CI_Upper = res$ci_upper
    ))
  }
}

# ------------------------------------------------------------------------------
# Create the Plots
# ------------------------------------------------------------------------------

# Rename sites for better labels
icc_plot_data <- icc_results %>%
  mutate(
    site_label = case_when(
      site == "MG" ~ "Mount Graham",
      site == "WM" ~ "White Mountains",
      TRUE ~ site
    ),
    # Clean parameter names
    param_clean = gsub("_", " ", parameter)
  )

# Plot
p_icc <- ggplot(icc_plot_data, aes(x = param_clean, y = ICC, color = site_label, shape = site_label)) +
  geom_point(position = position_dodge(width = 0.5), size = 6) +
  geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper),
                position = position_dodge(width = 0.5), width = 0.2, stroke = 1.5) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  scale_color_manual(values = c("Mount Graham" = "black", "White Mountains" = "black")) +
  scale_shape_manual(values = c("Mount Graham" = 19, "White Mountains" = 1)) + # 1 = open circle, 19 = filled circle
  labs(
    title = "ICC with Bootstrapped 95% Confidence Intervals",
    x = "Call Property",
    y = "ICC",
    color = "Population",
    shape = "Population"
  ) +
  theme_classic() +
  theme(
    legend.title = element_text(size = 14),
    legend.text  = element_text(size = 12),
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size = 18),
    axis.text.x  = element_text(size = 12, angle = 45, hjust = 1),
    axis.text.y  = element_text(size = 14)
  )

print(p_icc)
ggsave("figures/ICC_bootstrap_plot.png", plot = p_icc, width = 10, height = 6, dpi = 300)
 

```

## ICC table

```{r}
# Create a Clean Publication Table
# ------------------------------------------------------------------------------
library(knitr)

# Clean up the raw numbers and format the columns for a presentation-ready table
clean_icc_table <- icc_results %>%
  dplyr::mutate(
    parameter = str_replace_all(parameter, "_", " "), # Remove underscores
    parameter = str_to_title(parameter), # Capitalize
   
    # Format numbers strictly to 3 decimal places
    ICC = sprintf("%.3f", ICC),
    CI_Lower = sprintf("%.3f", CI_Lower),
    CI_Upper = sprintf("%.3f", CI_Upper),
   
    # Combine the Lower and Upper CIs into a single bracket format
    CI = paste0("[", CI_Lower, ", ", CI_Upper, "]")
  ) %>%
  dplyr::select(
    site,
    `Call Property` = parameter,
    ICC,
    CI
  ) %>%
  # Pivot wider to create separate ICC and CI columns for each population
  tidyr::pivot_wider(
    names_from = site,
    values_from = c(ICC, CI)
  ) %>%
  # Arrange columns to group Mount Graham (ICC, CI) and White Mountains (ICC, CI)
  dplyr::select(
    `Call Property`,
    `ICC` = ICC_MG,         # Select the automatically generated column ICC_MG
    `95% CI` = CI_MG,
    `ICC ` = ICC_WM,        # Added a space to make the column name technically unique in R
    `95% CI ` = CI_WM
  ) %>%
  dplyr::arrange(`Call Property`)

# (You can open this in Excel and copy/paste perfectly into Word)
write.csv(clean_icc_table, "tables/Clean_ICC_Bootstrap_Results.csv", row.names = FALSE)

# Print it using kable and kableExtra to create the overarching population titles!
library(kableExtra)
library(webshot2)

icc_formatted_table <- kable(clean_icc_table, align = "l") %>%
  add_header_above(c(" " = 1, "Mount Graham" = 2, "White Mountains" = 2)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

# Save the beautifully formatted table as an image in your figures folder
# Note: You may need to run `install.packages("webshot")` and `webshot::install_phantomjs()` once in your R console for this to work
save_kable(icc_formatted_table, "figures/Clean_ICC_Table.png")

# Still display it in the HTML document
icc_formatted_table

kable(clean_icc_table, align = "l") %>%
  add_header_above(c(" " = 1, "Mount Graham" = 2, "White Mountains" = 2)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

## PIC MG

```{r}
#PIC is a classic bioacoustic metric (Robisson et al. 1993) used to assess which call properties
#have the "potential" to encode individual identity. PIC = CV_among / CV_within.
#If PIC > 1, the parameter varies MORE between squirrels than within a single squirrel,
#meaning it could help distinguish individuals. Many older studies report PIC,
#so we calculate it here for direct comparison to other published literature.


# --- PIC + CVa + CVw + Partial Eta-Squared Calculation Function ---
# PIC = CVa / mean_CVw
# CVa (CV among) = SD of individual means / grand mean
# mean_CVw (mean CV within) = average of each individual's (SD / mean)
# Partial eta-squared = SS_among / (SS_among + SS_within) from a one-way ANOVA

calc_pic_full <- function(df, params_list) {
  # This function takes a call-level dataframe and a list of parameter names.
  # It calculates CVa, mean CVw, PIC, and partial eta-squared for each parameter.
 
  results <- tibble(
    parameter = character(),
    CVa = numeric(),
    mean_CVw = numeric(),
    PIC = numeric(),
    partial_eta_sq = numeric()
  )
 
  for (p in params_list) {
    # Step 1: For each squirrel, calculate the mean and SD of this parameter
    sq_stats <- df %>%
      group_by(sq_ID) %>%
      summarise(
        sq_mean = mean(.data[[p]], na.rm = TRUE),
        sq_sd   = sd(.data[[p]], na.rm = TRUE),
        .groups = "drop"
      ) %>%
      # CV for each individual = SD / mean
      mutate(sq_cv = sq_sd / sq_mean)
   
    # Step 2: CVa = SD of the individual means / grand mean
    grand_mean <- mean(sq_stats$sq_mean, na.rm = TRUE)
    sd_among   <- sd(sq_stats$sq_mean, na.rm = TRUE)
    cv_among   <- sd_among / grand_mean
   
    # Step 3: mean CVw = average of all individual CVs
    cv_within <- mean(sq_stats$sq_cv, na.rm = TRUE)
   
    # Step 4: PIC = CVa / mean_CVw
    pic_val <- cv_among / cv_within
   
    # Step 5: Partial eta-squared from a one-way ANOVA (parameter ~ sq_ID)
    # partial eta^2 = SS_between / (SS_between + SS_within)
    anova_formula <- as.formula(paste(p, "~ sq_ID"))
    anova_res <- aov(anova_formula, data = df)
    ss <- summary(anova_res)[[1]]["Sum Sq"]
    ss_among  <- ss[1, 1]  # SS for sq_ID (between groups)
    ss_within <- ss[2, 1]  # SS for Residuals (within groups)
    p_eta_sq  <- ss_among / (ss_among + ss_within)
   
    results <- bind_rows(results, tibble(
      parameter = p,
      CVa = cv_among,
      mean_CVw = cv_within,
      PIC = pic_val,
      partial_eta_sq = p_eta_sq
    ))
  }
 
  return(results)
}



## PIC MG


# Mount Graham PIC table
mg_pic_results <- calc_pic_full(call_avgs_mg_cleaned, params)

mg_pic_results <- mg_pic_results %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  arrange(desc(PIC))

mg_pic_results

write.csv(mg_pic_results, "tables/mg_pic_results.csv", row.names = FALSE)



## PIC WM


# White Mountains PIC table
wm_pic_results <- calc_pic_full(call_avgs_wm_cleaned, params)

wm_pic_results <- wm_pic_results %>%
  mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
  arrange(desc(PIC))

wm_pic_results

write.csv(wm_pic_results, "tables/wm_pic_results.csv", row.names = FALSE)



## Statistical Significance (ANOVA & FDR)

# Before relying on the ICC values, we must confirm that the variance among individuals is #statistically significant for each parameter. We run one-way ANOVAs for every acoustic #parameter by `sq_ID`. 
# Because we are running 17 tests per population, we must control for Type #I error inflation (false positives) by applying the Benjamini-Hochberg False Discovery Rate ##(FDR) correction to our p-values.


# ------------------------------------------------------------------------------
# ANOVA & FDR Correction for Individual Distinctiveness
# ------------------------------------------------------------------------------

# Function to run ANOVAs and extract p-values for all parameters
run_anovas <- function(df, params_list) {
  p_values <- c()
 
  for (p in params_list) {
    # Skip non-acoustic parameters
    if(p %in% c("no_notes", "call_duration")) next
   
    # Run a simple one-way ANOVA: parameter ~ sq_ID
    formula <- as.formula(paste(p, "~ sq_ID"))
    anova_res <- aov(formula, data = df)
   
    # Extract the p-value from the summary
    summary_anova <- summary(anova_res)
    p_val <- summary_anova[[1]][["Pr(>F)"]][1]
   
    p_values <- c(p_values, p_val)
  }
 
  # Return a dataframe with the raw p-values
  return(data.frame(
    parameter = params_list[!params_list %in% c("no_notes", "call_duration")],
    raw_p_value = p_values
  ))
}

# 1. Mount Graham Significance
mg_anova_results <- run_anovas(call_avgs_mg, params)
# Apply Benjamini-Hochberg (FDR) correction
mg_anova_results$fdr_adjusted_p <- p.adjust(mg_anova_results$raw_p_value, method = "BH")
# Mark significance (alpha = 0.05)
mg_anova_results$significant <- ifelse(mg_anova_results$fdr_adjusted_p < 0.05, "Yes", "No")

print("Mount Graham ANOVA & FDR Results:")
print(mg_anova_results)

# 2. White Mountains Significance
wm_anova_results <- run_anovas(call_avgs_wm, params)
# Apply Benjamini-Hochberg (FDR) correction
wm_anova_results$fdr_adjusted_p <- p.adjust(wm_anova_results$raw_p_value, method = "BH")
# Mark significance (alpha = 0.05)
wm_anova_results$significant <- ifelse(wm_anova_results$fdr_adjusted_p < 0.05, "Yes", "No")

print("White Mountains ANOVA & FDR Results:")
print(wm_anova_results)

# Save the significant results to your tables folder for reference
write.csv(mg_anova_results, "tables/MG_ANOVA_FDR_Results.csv", row.names = FALSE)
write.csv(wm_anova_results, "tables/WM_ANOVA_FDR_Results.csv", row.names = FALSE)
```

## PCA/Hs Beechers Stat

```{r}
## Principal Components Analysis

#Run separate Principal Components Analyses (PCA) for the Mount Graham and White Mountain squirrel populations to yield uncorrelated variables for Beecher's Information Statistic (Hs). Each is centered and scaled.


# ---------------------------------------------------------
# 1. MOUNT GRAHAM (MG) PCA
# ---------------------------------------------------------

# Filter specifically for Mount Graham squirrels and drop missing values
mg_pca_data <- call_avgs_mg_cleaned %>%
  dplyr::select(sq_ID, call_duration, note_rate, duration, break_s, disttomax,
         rms, energy, peaktopeak, peak_freq_mean, peak_ampl_mean,
         min_freq_mean, max_freq_mean, bandw_mean,
         quart_25_mean, quart_50_mean, quart_75_mean, entropy_mean) %>%
  drop_na()

# Run the PCA on just Mount Graham data (excluding the sq_ID column from the math)
mg.pc <- prcomp(~ call_duration + note_rate + duration + break_s + disttomax +
                  rms + energy + peaktopeak + peak_freq_mean + peak_ampl_mean +
                  min_freq_mean + max_freq_mean + bandw_mean +
                  quart_25_mean + quart_50_mean + quart_75_mean + entropy_mean,
                data=mg_pca_data, center=TRUE, scale=TRUE)

# Display a summary of the Mount Graham PCA results
summary(mg.pc)

# Extract and format the Loadings for MG
mg.load <- as.data.frame(mg.pc$rotation[,1:17])

# Extract importance stats (Standard Deviation & Variance) and append to Loadings
mg.pc.sum <- summary(mg.pc)[["importance"]]
mg.pc.sum <- rbind('Variance' = mg.pc.sum["Standard deviation",]^2, mg.pc.sum)
mg.load <- rbind(mg.load, mg.pc.sum[,1:17])

# Save Mount Graham Loadings to a CSV
write.csv(mg.load, file="tables/MG_Squirrel_PCA_Loadings.csv", row.names = TRUE)

# ---------------------------------------------------------
# MAKE MOUNT GRAHAM PCA TRANSFORMED DATA SET
# ---------------------------------------------------------

# For Mount Graham squirrels, we will retain the first 14 PCs which explain >99%
# of the variation (specifically, first 14 PCs explain 99.2%)
mg_scores_99 <- data.frame(mg.pc$x[, 1:14])
mg_scores_99$sq_ID <- mg_pca_data$sq_ID

# We will also save another data set retaining only PCs with SDs > 1
# (in this case it is the first 4 PCs). Together these PCs explain 77.8% of the variation.
mg_scores_sd1 <- data.frame(mg.pc$x[, 1:4])
mg_scores_sd1$sq_ID <- mg_pca_data$sq_ID

# Finally, save third data set that retains enough PC's to explain at least 95%
# of the variation (first 11 PCs explain 96.8%)
mg_scores_95 <- data.frame(mg.pc$x[, 1:11])
mg_scores_95$sq_ID <- mg_pca_data$sq_ID



# ---------------------------------------------------------
# 2. WHITE MOUNTAINS (WM) PCA
# ---------------------------------------------------------

# Filter specifically for White Mountain squirrels and drop missing values
wm_pca_data <- call_avgs_wm_cleaned %>%
  dplyr::select(sq_ID, call_duration, note_rate, duration, break_s, disttomax,
         rms, energy, peaktopeak, peak_freq_mean, peak_ampl_mean,
         min_freq_mean, max_freq_mean, bandw_mean,
         quart_25_mean, quart_50_mean, quart_75_mean, entropy_mean) %>%
  drop_na()

# Run the PCA on just White Mountain data (excluding the sq_ID column from the math)
wm.pc <- prcomp(~ call_duration + note_rate + duration + break_s + disttomax +
                  rms + energy + peaktopeak + peak_freq_mean + peak_ampl_mean +
                  min_freq_mean + max_freq_mean + bandw_mean +
                  quart_25_mean + quart_50_mean + quart_75_mean + entropy_mean,
                data=wm_pca_data, center=TRUE, scale=TRUE)

# Display a summary of the White Mountain PCA results
summary(wm.pc)

# Extract and format the Loadings for WM
wm.load <- as.data.frame(wm.pc$rotation[,1:17])

# Extract importance stats (Standard Deviation & Variance) and append to Loadings
wm.pc.sum <- summary(wm.pc)[["importance"]]
wm.pc.sum <- rbind('Variance' = wm.pc.sum["Standard deviation",]^2, wm.pc.sum)
wm.load <- rbind(wm.load, wm.pc.sum[,1:17])

# Save White Mountain Loadings to a CSV
write.csv(wm.load, file="tables/WM_Squirrel_PCA_Loadings.csv", row.names = TRUE)

# ---------------------------------------------------------
# MAKE WHITE MOUNTAINS PCA TRANSFORMED DATA SET
# ---------------------------------------------------------

# For White Mountains squirrels, we will retain the first 14 PCs which explain >99%
# of the variation (specifically, first 14 PCs explain 99.3%)
wm_scores_99 <- data.frame(wm.pc$x[, 1:14])
wm_scores_99$sq_ID <- wm_pca_data$sq_ID

# We will also save another data set retaining only PCs with SDs > 1
# (in this case it is the first 4 PCs). Together these PCs explain 80.0% of the variation.
wm_scores_sd1 <- data.frame(wm.pc$x[, 1:4])
wm_scores_sd1$sq_ID <- wm_pca_data$sq_ID

# Finally, save third data set that retains enough PC's to explain at least 95%
# of the variation (first 11 PCs explain 96.3%)
wm_scores_95 <- data.frame(wm.pc$x[, 1:11])
wm_scores_95$sq_ID <- wm_pca_data$sq_ID



## Hs (Beechers stat)

## Beecher's Hs Output

# Calculate Beecher's Information Statistic (Hs) using the three separate PCA datasets. We run ANOVAs for each Principal Component and sum the individual identity information (Hi) to yield the final Hs statistic for each population.

### Beecher's Hs: >99% Variation Dataset


# ---------------------------------------------------------
# MOUNT GRAHAM Hs (>99%)
# ---------------------------------------------------------

# Ensure sq_ID is a factor for the ANOVA
mg_scores_99$sq_ID <- as.factor(mg_scores_99$sq_ID)

# Create empty vector to hold Hi values
mg_H_99 <- vector()

# Loop through all PC columns (excluding the last column, which is sq_ID)
for(i in 1:(ncol(mg_scores_99)-1)){
  # Run ANOVA of PC score predicted by individual Squirrel ID
  fit2 <- aov(mg_scores_99[,i] ~ sq_ID, data=mg_scores_99)
  # Extract the F-statistic from the ANOVA summary
  F_stat <- summary(fit2)[[1]][["F value"]][1]
  # Get number of unique individuals (n)
  n <- length(levels(mg_scores_99$sq_ID))
 
  # Calculate Hi using Beecher's formula.
  # Note: if F_stat is less than 1, we set it to 1 to prevent getting NaNs from log2 of negative numbers
  F_safe <- ifelse(F_stat < 1, 1, F_stat)
  Hi <- log2(sqrt((F_safe + n - 1) / n))
  mg_H_99[i] <- Hi
}

# Sum the Hi values to get total Hs
mg_Hs_99 <- sum(mg_H_99)
print(paste("Mount Graham Hs (>99% Var) =", round(mg_Hs_99, 2)))


# ---------------------------------------------------------
# WHITE MOUNTAINS Hs (>99%)
# ---------------------------------------------------------

wm_scores_99$sq_ID <- as.factor(wm_scores_99$sq_ID)
wm_H_99 <- vector()

for(i in 1:(ncol(wm_scores_99)-1)){
  fit2 <- aov(wm_scores_99[,i] ~ sq_ID, data=wm_scores_99)
  F_stat <- summary(fit2)[[1]][["F value"]][1]
  n <- length(levels(wm_scores_99$sq_ID))
 
  F_safe <- ifelse(F_stat < 1, 1, F_stat)
  Hi <- log2(sqrt((F_safe + n - 1) / n))
  wm_H_99[i] <- Hi
}

wm_Hs_99 <- sum(wm_H_99)
print(paste("White Mountains Hs (>99% Var) =", round(wm_Hs_99, 2)))



### Beecher's Hs: Kaiser's Rule (SD > 1) Dataset


# ---------------------------------------------------------
# MOUNT GRAHAM Hs (SD > 1)
# ---------------------------------------------------------

mg_scores_sd1$sq_ID <- as.factor(mg_scores_sd1$sq_ID)
mg_H_sd1 <- vector()

for(i in 1:(ncol(mg_scores_sd1)-1)){
  fit2 <- aov(mg_scores_sd1[,i] ~ sq_ID, data=mg_scores_sd1)
  F_stat <- summary(fit2)[[1]][["F value"]][1]
  n <- length(levels(mg_scores_sd1$sq_ID))
 
  F_safe <- ifelse(F_stat < 1, 1, F_stat)
  Hi <- log2(sqrt((F_safe + n - 1) / n))
  mg_H_sd1[i] <- Hi
}

mg_Hs_sd1 <- sum(mg_H_sd1)
print(paste("Mount Graham Hs (SD > 1) =", round(mg_Hs_sd1, 2)))


# ---------------------------------------------------------
# WHITE MOUNTAINS Hs (SD > 1)
# ---------------------------------------------------------

wm_scores_sd1$sq_ID <- as.factor(wm_scores_sd1$sq_ID)
wm_H_sd1 <- vector()

for(i in 1:(ncol(wm_scores_sd1)-1)){
  fit2 <- aov(wm_scores_sd1[,i] ~ sq_ID, data=wm_scores_sd1)
  F_stat <- summary(fit2)[[1]][["F value"]][1]
  n <- length(levels(wm_scores_sd1$sq_ID))
 
  F_safe <- ifelse(F_stat < 1, 1, F_stat)
  Hi <- log2(sqrt((F_safe + n - 1) / n))
  wm_H_sd1[i] <- Hi
}

wm_Hs_sd1 <- sum(wm_H_sd1)
print(paste("White Mountains Hs (SD > 1) =", round(wm_Hs_sd1, 2)))



### Beecher's Hs: >95% Variation Dataset


# ---------------------------------------------------------
# MOUNT GRAHAM Hs (>95%)
# ---------------------------------------------------------

mg_scores_95$sq_ID <- as.factor(mg_scores_95$sq_ID)
mg_H_95 <- vector()

for(i in 1:(ncol(mg_scores_95)-1)){
  fit2 <- aov(mg_scores_95[,i] ~ sq_ID, data=mg_scores_95)
  F_stat <- summary(fit2)[[1]][["F value"]][1]
  n <- length(levels(mg_scores_95$sq_ID))
 
  F_safe <- ifelse(F_stat < 1, 1, F_stat)
  Hi <- log2(sqrt((F_safe + n - 1) / n))
  mg_H_95[i] <- Hi
}

mg_Hs_95 <- sum(mg_H_95)
print(paste("Mount Graham Hs (>95% Var) =", round(mg_Hs_95, 2)))


# ---------------------------------------------------------
# WHITE MOUNTAINS Hs (>95%)
# ---------------------------------------------------------

wm_scores_95$sq_ID <- as.factor(wm_scores_95$sq_ID)
wm_H_95 <- vector()

for(i in 1:(ncol(wm_scores_95)-1)){
  fit2 <- aov(wm_scores_95[,i] ~ sq_ID, data=wm_scores_95)
  F_stat <- summary(fit2)[[1]][["F value"]][1]
  n <- length(levels(wm_scores_95$sq_ID))
 
  F_safe <- ifelse(F_stat < 1, 1, F_stat)
  Hi <- log2(sqrt((F_safe + n - 1) / n))
  wm_H_95[i] <- Hi
}

wm_Hs_95 <- sum(wm_H_95)
print(paste("White Mountains Hs (>95% Var) =", round(wm_Hs_95, 2)))
```

## Linear Discriminat Analysis LDA

```         
```

```{r}
## Linear Discriminant Analysis (LDA)

#Run a Linear Discriminant Analysis (LDA) on the PCA scores to test classification success.
#First, we run the LDA on the full populations for all three datasets to extract singular values, scaling, and overall classification success via Leave-One-Out Cross Validation (CV=TRUE). Then we run it on groups of 5 individuals to control for baseline probability differences caused by unequal population sizes.

{warning=FALSE}
# Load the MASS package which contains the lda() function
library(MASS)

# =========================================================
# MOUNT GRAHAM (MG) FULL LDA
# =========================================================

# ---------------------------------------------------------
# MG: >99% Variation Dataset (14 PCs)
# ---------------------------------------------------------
mg.dfa_99 <- lda(sq_ID ~., mg_scores_99)

print("MG (>99% Var) Scaling of PCs on linear discriminants (first 8):")
round(as.data.frame(mg.dfa_99$scaling)[1:8], 2)

print("MG (>99% Var) Singular values of linear discriminants:")
mg.dfa_99$svd

# Make data frame of scaling and singular values plus proportion of trace
mg.dfa_99.scale <- rbind(as.data.frame(mg.dfa_99$scaling), 'svd'=mg.dfa_99$svd,
                         'prop_trace'=mg.dfa_99$svd^2/sum(mg.dfa_99$svd^2),
                         'cumprop_trace'=cumsum(mg.dfa_99$svd^2)/sum(mg.dfa_99$svd^2))

# CV=TRUE tells it to do leave-one-out cross validation
mg.dfa_99CV <- lda(sq_ID ~., mg_scores_99, CV=TRUE)
match <- ifelse(mg.dfa_99CV$class==mg_scores_99$sq_ID, 1, 0)
score <- sum(match)/length(mg.dfa_99CV$class)*100
print(paste("MG (>99% Var) % classification success =", round(score, 2)))

# Create data frame of classification for each call
mg.class_99 <- data.frame(mg_scores_99$sq_ID)
mg.class_99$class <- mg.dfa_99CV$class
colnames(mg.class_99) <- c("ID", "class")
mg.class_99$score <- ifelse(mg.class_99$ID == mg.class_99$class, 1, 0)
mg.class_99$species <- "MG"

# ---------------------------------------------------------
# MG: SD > 1 Dataset (4 PCs)
# ---------------------------------------------------------
mg.dfa_sd1 <- lda(sq_ID ~., mg_scores_sd1)

print("MG (SD > 1) Scaling of PCs on linear discriminants:")
round(as.data.frame(mg.dfa_sd1$scaling), 2)

print("MG (SD > 1) Singular values of linear discriminants:")
mg.dfa_sd1$svd

mg.dfa_sd1.scale <- rbind(as.data.frame(mg.dfa_sd1$scaling), 'svd'=mg.dfa_sd1$svd,
                         'prop_trace'=mg.dfa_sd1$svd^2/sum(mg.dfa_sd1$svd^2),
                         'cumprop_trace'=cumsum(mg.dfa_sd1$svd^2)/sum(mg.dfa_sd1$svd^2))

mg.dfa_sd1CV <- lda(sq_ID ~., mg_scores_sd1, CV=TRUE)
match <- ifelse(mg.dfa_sd1CV$class==mg_scores_sd1$sq_ID, 1, 0)
score <- sum(match)/length(mg.dfa_sd1CV$class)*100
print(paste("MG (SD > 1) % classification success =", round(score, 2)))

mg.class_sd1 <- data.frame(mg_scores_sd1$sq_ID)
mg.class_sd1$class <- mg.dfa_sd1CV$class
colnames(mg.class_sd1) <- c("ID", "class")
mg.class_sd1$score <- ifelse(mg.class_sd1$ID == mg.class_sd1$class, 1, 0)
mg.class_sd1$species <- "MG"

# ---------------------------------------------------------
# MG: >95% Variation Dataset (11 PCs)
# ---------------------------------------------------------
mg.dfa_95 <- lda(sq_ID ~., mg_scores_95)

print("MG (>95% Var) Scaling of PCs on linear discriminants (first 8):")
round(as.data.frame(mg.dfa_95$scaling)[1:8], 2)

print("MG (>95% Var) Singular values of linear discriminants:")
mg.dfa_95$svd

mg.dfa_95.scale <- rbind(as.data.frame(mg.dfa_95$scaling), 'svd'=mg.dfa_95$svd,
                         'prop_trace'=mg.dfa_95$svd^2/sum(mg.dfa_95$svd^2),
                         'cumprop_trace'=cumsum(mg.dfa_95$svd^2)/sum(mg.dfa_95$svd^2))

mg.dfa_95CV <- lda(sq_ID ~., mg_scores_95, CV=TRUE)
match <- ifelse(mg.dfa_95CV$class==mg_scores_95$sq_ID, 1, 0)
score <- sum(match)/length(mg.dfa_95CV$class)*100
print(paste("MG (>95% Var) % classification success =", round(score, 2)))

mg.class_95 <- data.frame(mg_scores_95$sq_ID)
mg.class_95$class <- mg.dfa_95CV$class
colnames(mg.class_95) <- c("ID", "class")
mg.class_95$score <- ifelse(mg.class_95$ID == mg.class_95$class, 1, 0)
mg.class_95$species <- "MG"


# =========================================================
# WHITE MOUNTAINS (WM) FULL LDA
# =========================================================

# ---------------------------------------------------------
# WM: >99% Variation Dataset (14 PCs)
# ---------------------------------------------------------
wm.dfa_99 <- lda(sq_ID ~., wm_scores_99)

print("WM (>99% Var) Scaling of PCs on linear discriminants (first 8):")
round(as.data.frame(wm.dfa_99$scaling)[1:8], 2)

print("WM (>99% Var) Singular values of linear discriminants:")
wm.dfa_99$svd

wm.dfa_99.scale <- rbind(as.data.frame(wm.dfa_99$scaling), 'svd'=wm.dfa_99$svd,
                         'prop_trace'=wm.dfa_99$svd^2/sum(wm.dfa_99$svd^2),
                         'cumprop_trace'=cumsum(wm.dfa_99$svd^2)/sum(wm.dfa_99$svd^2))

wm.dfa_99CV <- lda(sq_ID ~., wm_scores_99, CV=TRUE)
match <- ifelse(wm.dfa_99CV$class==wm_scores_99$sq_ID, 1, 0)
score <- sum(match)/length(wm.dfa_99CV$class)*100
print(paste("WM (>99% Var) % classification success =", round(score, 2)))

wm.class_99 <- data.frame(wm_scores_99$sq_ID)
wm.class_99$class <- wm.dfa_99CV$class
colnames(wm.class_99) <- c("ID", "class")
wm.class_99$score <- ifelse(wm.class_99$ID == wm.class_99$class, 1, 0)
wm.class_99$species <- "WM"

# ---------------------------------------------------------
# WM: SD > 1 Dataset (4 PCs)
# ---------------------------------------------------------
wm.dfa_sd1 <- lda(sq_ID ~., wm_scores_sd1)

print("WM (SD > 1) Scaling of PCs on linear discriminants:")
round(as.data.frame(wm.dfa_sd1$scaling), 2)

print("WM (SD > 1) Singular values of linear discriminants:")
wm.dfa_sd1$svd

wm.dfa_sd1.scale <- rbind(as.data.frame(wm.dfa_sd1$scaling), 'svd'=wm.dfa_sd1$svd,
                         'prop_trace'=wm.dfa_sd1$svd^2/sum(wm.dfa_sd1$svd^2),
                         'cumprop_trace'=cumsum(wm.dfa_sd1$svd^2)/sum(wm.dfa_sd1$svd^2))

wm.dfa_sd1CV <- lda(sq_ID ~., wm_scores_sd1, CV=TRUE)
match <- ifelse(wm.dfa_sd1CV$class==wm_scores_sd1$sq_ID, 1, 0)
score <- sum(match)/length(wm.dfa_sd1CV$class)*100
print(paste("WM (SD > 1) % classification success =", round(score, 2)))

wm.class_sd1 <- data.frame(wm_scores_sd1$sq_ID)
wm.class_sd1$class <- wm.dfa_sd1CV$class
colnames(wm.class_sd1) <- c("ID", "class")
wm.class_sd1$score <- ifelse(wm.class_sd1$ID == wm.class_sd1$class, 1, 0)
wm.class_sd1$species <- "WM"

# ---------------------------------------------------------
# WM: >95% Variation Dataset (11 PCs)
# ---------------------------------------------------------
wm.dfa_95 <- lda(sq_ID ~., wm_scores_95)

print("WM (>95% Var) Scaling of PCs on linear discriminants (first 8):")
round(as.data.frame(wm.dfa_95$scaling)[1:8], 2)

print("WM (>95% Var) Singular values of linear discriminants:")
wm.dfa_95$svd

wm.dfa_95.scale <- rbind(as.data.frame(wm.dfa_95$scaling), 'svd'=wm.dfa_95$svd,
                         'prop_trace'=wm.dfa_95$svd^2/sum(wm.dfa_95$svd^2),
                         'cumprop_trace'=cumsum(wm.dfa_95$svd^2)/sum(wm.dfa_95$svd^2))

wm.dfa_95CV <- lda(sq_ID ~., wm_scores_95, CV=TRUE)
match <- ifelse(wm.dfa_95CV$class==wm_scores_95$sq_ID, 1, 0)
score <- sum(match)/length(wm.dfa_95CV$class)*100
print(paste("WM (>95% Var) % classification success =", round(score, 2)))

wm.class_95 <- data.frame(wm_scores_95$sq_ID)
wm.class_95$class <- wm.dfa_95CV$class
colnames(wm.class_95) <- c("ID", "class")
wm.class_95$score <- ifelse(wm.class_95$ID == wm.class_95$class, 1, 0)
wm.class_95$species <- "WM"


# =========================================================
# LDA ON GROUPS OF 5 INDIVIDUALS
# =========================================================
# Run leave-one-out cross validation on 100 randomly selected groups of five
# individuals, return mean and range of classification success

# ---------------------------------------------------------
# MOUNT GRAHAM LDA (Groups of 5)
# ---------------------------------------------------------

### MG: >99% Variation Dataset (14 PCs)
mg_correct_99 <- vector()
for (i in 1:100){
  IDs <- sample(levels(mg_scores_99$sq_ID), 5, replace=FALSE)
  pcscores_sample <- mg_scores_99[mg_scores_99$sq_ID %in% IDs,]
  pcscores_sample$sq_ID <- factor(pcscores_sample$sq_ID)
 
  dfa_sample <- lda(sq_ID ~., pcscores_sample, CV=TRUE)
 
  match_sample <- ifelse(dfa_sample$class == pcscores_sample$sq_ID, 1, 0)
  score <- sum(match_sample) / length(dfa_sample$class) * 100
  mg_correct_99[i] <- score
}
print(paste("MG (>99% Var - Groups of 5) mean % success =", round(mean(mg_correct_99), 2)))
print(paste("MG (>99% Var - Groups of 5) range =", round(min(mg_correct_99), 2), "-", round(max(mg_correct_99), 2)))

### MG: SD > 1 Dataset (4 PCs)
mg_correct_sd1 <- vector()
for (i in 1:100){
  IDs <- sample(levels(mg_scores_sd1$sq_ID), 5, replace=FALSE)
  pcscores_sample <- mg_scores_sd1[mg_scores_sd1$sq_ID %in% IDs,]
  pcscores_sample$sq_ID <- factor(pcscores_sample$sq_ID)
 
  dfa_sample <- lda(sq_ID ~., pcscores_sample, CV=TRUE)
 
  match_sample <- ifelse(dfa_sample$class == pcscores_sample$sq_ID, 1, 0)
  score <- sum(match_sample) / length(dfa_sample$class) * 100
  mg_correct_sd1[i] <- score
}
print(paste("MG (SD > 1 - Groups of 5) mean % success =", round(mean(mg_correct_sd1), 2)))
print(paste("MG (SD > 1 - Groups of 5) range =", round(min(mg_correct_sd1), 2), "-", round(max(mg_correct_sd1), 2)))

### MG: >95% Variation Dataset (11 PCs)
mg_correct_95 <- vector()
for (i in 1:100){
  IDs <- sample(levels(mg_scores_95$sq_ID), 5, replace=FALSE)
  pcscores_sample <- mg_scores_95[mg_scores_95$sq_ID %in% IDs,]
  pcscores_sample$sq_ID <- factor(pcscores_sample$sq_ID)
 
  dfa_sample <- lda(sq_ID ~., pcscores_sample, CV=TRUE)
 
  match_sample <- ifelse(dfa_sample$class == pcscores_sample$sq_ID, 1, 0)
  score <- sum(match_sample) / length(dfa_sample$class) * 100
  mg_correct_95[i] <- score
}
print(paste("MG (>95% Var - Groups of 5) mean % success =", round(mean(mg_correct_95), 2)))
print(paste("MG (>95% Var - Groups of 5) range =", round(min(mg_correct_95), 2), "-", round(max(mg_correct_95), 2)))


# ---------------------------------------------------------
# WHITE MOUNTAINS LDA (Groups of 5)
# ---------------------------------------------------------

### WM: >99% Variation Dataset (14 PCs)
wm_correct_99 <- vector()
for (i in 1:100){
  IDs <- sample(levels(wm_scores_99$sq_ID), 5, replace=FALSE)
  pcscores_sample <- wm_scores_99[wm_scores_99$sq_ID %in% IDs,]
  pcscores_sample$sq_ID <- factor(pcscores_sample$sq_ID)
 
  dfa_sample <- lda(sq_ID ~., pcscores_sample, CV=TRUE)
 
  match_sample <- ifelse(dfa_sample$class == pcscores_sample$sq_ID, 1, 0)
  score <- sum(match_sample) / length(dfa_sample$class) * 100
  wm_correct_99[i] <- score
}
print(paste("WM (>99% Var - Groups of 5) mean % success =", round(mean(wm_correct_99), 2)))
print(paste("WM (>99% Var - Groups of 5) range =", round(min(wm_correct_99), 2), "-", round(max(wm_correct_99), 2)))

### WM: SD > 1 Dataset (4 PCs)
wm_correct_sd1 <- vector()
for (i in 1:100){
  IDs <- sample(levels(wm_scores_sd1$sq_ID), 5, replace=FALSE)
  pcscores_sample <- wm_scores_sd1[wm_scores_sd1$sq_ID %in% IDs,]
  pcscores_sample$sq_ID <- factor(pcscores_sample$sq_ID)
 
  dfa_sample <- lda(sq_ID ~., pcscores_sample, CV=TRUE)
 
  match_sample <- ifelse(dfa_sample$class == pcscores_sample$sq_ID, 1, 0)
  score <- sum(match_sample) / length(dfa_sample$class) * 100
  wm_correct_sd1[i] <- score
}
print(paste("WM (SD > 1 - Groups of 5) mean % success =", round(mean(wm_correct_sd1), 2)))
print(paste("WM (SD > 1 - Groups of 5) range =", round(min(wm_correct_sd1), 2), "-", round(max(wm_correct_sd1), 2)))

### WM: >95% Variation Dataset (11 PCs)
wm_correct_95 <- vector()
for (i in 1:100){
  IDs <- sample(levels(wm_scores_95$sq_ID), 5, replace=FALSE)
  pcscores_sample <- wm_scores_95[wm_scores_95$sq_ID %in% IDs,]
  pcscores_sample$sq_ID <- factor(pcscores_sample$sq_ID)
 
  dfa_sample <- lda(sq_ID ~., pcscores_sample, CV=TRUE)
 
  match_sample <- ifelse(dfa_sample$class == pcscores_sample$sq_ID, 1, 0)
  score <- sum(match_sample) / length(dfa_sample$class) * 100
  wm_correct_95[i] <- score
}
print(paste("WM (>95% Var - Groups of 5) mean % success =", round(mean(wm_correct_95), 2)))
print(paste("WM (>95% Var - Groups of 5) range =", round(min(wm_correct_95), 2), "-", round(max(wm_correct_95), 2)))

```

## LDA Plots

```{r}
# LDA Classification Plots

### Mount Graham

# Select 5 random individuals and plot in 1st 2 dimensions of linear discriminant analysis. Re-run to resample and plot 5 new individuals.


library(ggplot2)
library(lme4)
library(lmtest)
library(wesanderson)
# ---------------------------------------------------------
# MOUNT GRAHAM LDA Plot (Faceted 4-Iteration Resampling)
# ---------------------------------------------------------
# Run 4 iterations of resampling 5 individuals and bind the results together

# Initialize empty lists to store iteration data
mg_dfa_list <- list()
mg_centroids_list <- list()

for (i in 1:4) {
  # Resample and predict
  IDs_mg <- sample(levels(mg_scores_99$sq_ID), 5, replace=FALSE)
  pcscores_sample_mg <- mg_scores_99[mg_scores_99$sq_ID %in% IDs_mg,]
  pcscores_sample_mg$sq_ID <- factor(pcscores_sample_mg$sq_ID)
 
  dfa_sample_mg <- lda(sq_ID ~., pcscores_sample_mg)
  p.dfa_mg <- predict(object = dfa_sample_mg, newdata = pcscores_sample_mg)

  # Save results and add an iteration label for faceting
  dfa_results <- data.frame(sq_ID = pcscores_sample_mg[,"sq_ID"], lda = p.dfa_mg$x)
  dfa_results$Iteration <- paste("Iteration", i)
 
  # create a local 1-5 index so the 5-color palette scales correctly out of 20 possible IDs
  dfa_results$Local_ID <- factor(as.numeric(factor(dfa_results$sq_ID)))
 
  centroids <- aggregate(cbind(lda.LD1, lda.LD2) ~ Local_ID, dfa_results, mean)
  centroids$Iteration <- paste("Iteration", i)
 
  mg_dfa_list[[i]] <- dfa_results
  mg_centroids_list[[i]] <- centroids
}

# Bind lists into full data frames
mg_dfa_all <- do.call(rbind, mg_dfa_list)
mg_centroids_all <- do.call(rbind, mg_centroids_list)

# Generate Faceted Plot
ggplot(mg_dfa_all, aes(x = lda.LD1, y = lda.LD2, colour=Local_ID)) +
  geom_point(size = 0.5) +
  stat_ellipse(level=0.95, size=0.4) +
  geom_point(data=mg_centroids_all, aes(fill=Local_ID), size=2, pch=21, colour="black") +
  scale_colour_manual(values = wes_palette("Darjeeling1", n = 5, type = "discrete")) +
  scale_fill_manual(values = wes_palette("Darjeeling1", n = 5, type = "discrete")) +
  facet_wrap(~Iteration, scales = "free") +
  theme_bw() +
  ggtitle("Mount Graham") +
  xlab("Linear Discriminant 1") +
  ylab("Linear Discriminant 2") +
  theme(legend.position = "none", panel.grid = element_blank(),
        plot.title = element_text(face="bold", hjust=0.5), axis.text = element_text(size=6),
        strip.text = element_blank(), strip.background = element_blank())

ggsave("figures/Mount_Graham_LDA_Faceted.png", width = 8, height = 6)


### White Mountains

# Select 5 random individuals and plot in 1st 2 dimensions of linear discriminant analysis. Re-run to resample and plot 5 new individuals.


# ---------------------------------------------------------
# WHITE MOUNTAINS LDA Plot (Faceted 4-Iteration Resampling)
# ---------------------------------------------------------
# Run 4 iterations of resampling 5 individuals and bind the results together

# Initialize empty lists to store iteration data
wm_dfa_list <- list()
wm_centroids_list <- list()

for (i in 1:4) {
  # Resample and predict
  IDs_wm <- sample(levels(wm_scores_99$sq_ID), 5, replace=FALSE)
  pcscores_sample_wm <- wm_scores_99[wm_scores_99$sq_ID %in% IDs_wm,]
  pcscores_sample_wm$sq_ID <- factor(pcscores_sample_wm$sq_ID)
 
  dfa_sample_wm <- lda(sq_ID ~., pcscores_sample_wm)
  p.dfa_wm <- predict(object = dfa_sample_wm, newdata = pcscores_sample_wm)

  # Save results and add an iteration label for faceting
  dfa_results <- data.frame(sq_ID = pcscores_sample_wm[,"sq_ID"], lda = p.dfa_wm$x)
  dfa_results$Iteration <- paste("Iteration", i)
 
  # create a local 1-5 index so the 5-color palette scales correctly out of 20 possible IDs
  dfa_results$Local_ID <- factor(as.numeric(factor(dfa_results$sq_ID)))
 
  centroids <- aggregate(cbind(lda.LD1, lda.LD2) ~ Local_ID, dfa_results, mean)
  centroids$Iteration <- paste("Iteration", i)
 
  wm_dfa_list[[i]] <- dfa_results
  wm_centroids_list[[i]] <- centroids
}

# Bind lists into full data frames
wm_dfa_all <- do.call(rbind, wm_dfa_list)
wm_centroids_all <- do.call(rbind, wm_centroids_list)

# Generate Faceted Plot
ggplot(wm_dfa_all, aes(x = lda.LD1, y = lda.LD2, colour=Local_ID)) +
  geom_point(size = 0.5) +
  stat_ellipse(level=0.95, size=0.4) +
  geom_point(data=wm_centroids_all, aes(fill=Local_ID), size=2, pch=21, colour="black") +
  scale_colour_manual(values = wes_palette("Darjeeling1", n = 5, type = "discrete")) +
  scale_fill_manual(values = wes_palette("Darjeeling1", n = 5, type = "discrete")) +
  facet_wrap(~Iteration, scales = "free") +
  theme_bw() +
  ggtitle("White Mountains") +
  xlab("Linear Discriminant 1") +
  ylab("Linear Discriminant 2") +
  theme(legend.position = "none", panel.grid = element_blank(),
        plot.title = element_text(face="bold", hjust=0.5), axis.text = element_text(size=6),
        strip.text = element_blank(), strip.background = element_blank())

ggsave("figures/White_Mountains_LDA_Faceted.png", width = 8, height = 6)


### Classification success species comparison plot

# Mean and range of classification success from 100 randomly selected groups of 5 individuals for each species


mg_correct_df <- as.data.frame(mg_correct_99)
mg_correct_df$species <- "Mount Graham"
colnames(mg_correct_df) <- c("correctsample", "species")

wm_correct_df <- as.data.frame(wm_correct_99)
wm_correct_df$species <- "White Mountains"
colnames(wm_correct_df) <- c("correctsample", "species")

species.correctsample <- rbind(mg_correct_df, wm_correct_df)

ggplot(species.correctsample, aes(x=species, y=correctsample, fill=species)) +
  stat_summary(fun.min = min, fun.max = max, geom="errorbar", width=0) +
  stat_summary(fun = mean, geom="point", pch=21, size=5) +
  scale_y_continuous(limits=c(0,100), breaks=seq(0,100,20)) +
  geom_hline(yintercept = 20, lty=2) +
  ylab("Call classification success (%)") +
  theme_bw() +
  theme(panel.grid = element_blank(), axis.title.x = element_blank(), legend.position="none")
ggsave("figures/Classification_success_comparison.png", width = 6, height = 5)

# Classification success species comparison

# In order to statistically compare the identity info in Mount Graham and White Mountains calls, the classification success of each call (in the full data set) is coded as 0 or 1, based on whether or not the call was assigned to the correct individual.

# Then we bind together the two data sets and run a logistic regression with species as the main effect and individual as a random effect, and ask if there is any species difference in the classification success

# Return summary of model and likelihood ratio test to test significance of species

### Species Comparison from LDA on PC Scores


#Bind together classification data sets
classdata_99 <- rbind(mg.class_99, wm.class_99)

m1 <- glmer(score ~ species + (1|ID), data=classdata_99, family="binomial")
lrtest(m1, update(m1, ~. -species))


### Species Comparison from LDA on PC scores from PCs with SDs > 1


#Bind together classification data sets
classdata_sd1 <- rbind(mg.class_sd1, wm.class_sd1)

m2 <- glmer(score ~ species + (1|ID), data=classdata_sd1, family="binomial")
lrtest(m2, update(m2, ~. -species))


### Species Comparison from LDA on PC scores from PCs that explain 95% of variance


#Bind together classification data sets
classdata_95 <- rbind(mg.class_95, wm.class_95)

m3 <- glmer(score ~ species + (1|ID), data=classdata_95, family="binomial")
lrtest(m3, update(m3, ~. -species))
```
