---
title: "First pass analysis"
author: "Nikki Reck"
format: html
editor: visual
---

## Background

This is my first pass analysis for my data on red squirrel vocal distinctiveness. I will be cleaning the data so that it is compatible with R (renaming columns, and creating new columns). Then I will take averages of each call from each animal, then take averages of all calls per animal. Then I will look at how the two populations may or may not differ.

## Load Data

Load in packages and the data sets I will be using.

```{r}
library(lme4)
library(effectsize)
library(tidyverse)


  sq_data <- read_csv("data_raw/combined_sq_data6.csv")
view(sq_data)

detections <- read_csv("data_raw/31dec2025_detections9.csv")

# create a new column for call number, gsub will cut out the file name so that i just have sq id midden date time.

# Clean break_s: "_" = missing, then make numeric
detections <- detections %>%
  rename(break_s = break) %>%
  mutate(
    break_s = as.character(break_s),
    break_s = na_if(break_s, "")
  )




```

sq_rattle_data %\>%

filter(site == "MG") %\>% \# only Mount Graham data

summarise(

n_squirrels = n_distinct(sq_ID) \# number of unique squirrels

)

```{r}

# load data
detections <- read_csv("data_raw/31dec2025_detections9.csv")
combined   <- read_csv("data_raw/combined_sq_data6.csv")

# created a unique call id number so that the note number doesnt collapse
detections$uniqueid <- sub(".*\\\\(.*)\\.wav$", "\\1", detections$filename) 


# combine data. Try the merge function merge(x, y, by.x = "whatever column", "column, by .y = specify although x and although y, (It will only merge things that are the same what my columns are)

sq_rattle_data<- merge(detections, combined, by.x = "midden", by.y = "midden", all.x = TRUE, all.y = FALSE) %>% 
  arrange(midden, uniqueid, rattle_no, note_number)

# ---- combined: build uniqueid ----
combined <- combined %>%
  mutate(
    # TOD as 4-digit text (works whether TOD is numeric or already "0645")
    TOD = str_pad(as.character(TOD), width = 4, pad = "0"),

    # parse date no matter whether it's "06/20/25", "06/20/2025", or "06202025"
    date_parsed = case_when(
      str_detect(as.character(date), "^\\d{2}/\\d{2}/\\d{2}$") ~ as.Date(as.character(date), format = "%m/%d/%y"),
      str_detect(as.character(date), "^\\d{2}/\\d{2}/\\d{4}$") ~ as.Date(as.character(date), format = "%m/%d/%Y"),
      str_detect(as.character(date), "^\\d{8}$")               ~ as.Date(as.character(date), format = "%m%d%Y"),
      TRUE                                                     ~ NA_Date_
    ),

    uniqueid = paste0(midden, "_", format(date_parsed, "%m%d%Y"), "_", TOD)
  )

# ---- detections: extract uniqueid from filename ----
detections <- detections %>%
  mutate(
    uniqueid = filename %>%
      str_replace("^.*[\\\\/]", "") %>%   # drop any path (works for \ and /)
      str_remove("\\.wav$") %>%          # drop extension
      if_else(str_count(., "_") > 2, str_replace(., "^[^_]*_", ""), .)  # drop leading prefix if present
  )

# ---- merge ----
sq_rattle_data <- merge(detections, combined, by.x = "uniqueid", by.y = "uniqueid")




#### quick check ssee if all squirrels merged correctly #####

# mt graham squirrels

#  sq data BEFORE i merged data frames
combined %>%
  filter(site == "MG") %>%
  summarise(
    sq_IDs = list(sort(unique(sq_ID)))
  )
# 19 sqs

# sq data AFTER I merge data frames
sq_rattle_data %>%
  filter(site == "MG") %>%
  summarise(
    sq_IDs = list(sort(unique(sq_ID)))
  )
# 19sqs


################

# White mountain squirrels

#  sq data BEOFRE i merged data frames
combined %>%
  filter(site == "WM") %>%
  summarise(
    sq_IDs = list(sort(unique(sq_ID)))
  )
# 21 sqs



# sq data AFTER I merge data frames
sq_rattle_data %>%
  filter(site == "WM") %>%
  summarise(
    sq_IDs = list(sort(unique(sq_ID)))
  )
# 21 sqs

```

## Distance bias check on parameters

```{r}

# Calculate note rate
sq_rattle_data <- sq_rattle_data %>%
  group_by(uniqueid, rattle_no) %>% 
  mutate(
    note_rate = n() / first(call_duration)
  ) %>%
  ungroup()


params <- c(
  "call_duration",
  "note_number",
  "note_rate",
  "duration",
  "break_s",
  "disttomax",
  "rms",
  "energy",
  "peaktopeak",
  "peak_freq_mean",
  "peak_ampl_mean",
  "min_freq_mean",
  "max_freq_mean",
  "bandw_mean",
  "quart_25_mean",
  "quart_50_mean",
  "quart_75_mean",
  "entropy_mean"
)


dir.create("figures/distance_bias", recursive = TRUE, showWarnings = FALSE)

for (p in params) {

  g <- ggplot(
    sq_rattle_data,
    aes(x = distance, y = .data[[p]], color = site)
  ) +
    geom_point(na.rm = TRUE, alpha = 0.7) +
    geom_smooth(
      method = "lm",
      se = FALSE,
      na.rm = TRUE
    ) +
    scale_color_manual(
      values = c(
        "Mount Graham" = "#009E73",   # bluish-green
        "White Mountains" = "#CC79A7" # purple
      )
    ) +
    labs(
      title = paste("Distance bias check:", p),
      x = "Distance recorded (m)",
      y = p,
      color = "Population"
    ) +
    theme_bw() +
    theme(
      plot.title = element_text(face = "bold"),
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10),
      legend.position = "top"
    )

  ggsave(
    filename = file.path(
      "figures/distance_bias",
      paste0("distance_bias_", p, ".png")
    ),
    plot = g,
    width = 7,
    height = 5,
    dpi = 300
  )
}

```

## Call level Averages

Taking the averages of each of the calls per animal. Fix any oddities.

```{r}

# Call-level averages by sq_ID (one row per call)

call_avgs <- sq_rattle_data %>%
  group_by(sq_ID, site, filename) %>%
  summarise(
    no_notes = n(),
    call_duration = first(call_duration),
    note_rate = no_notes / call_duration,

    duration = mean(duration, na.rm = TRUE),
    break_s = mean(break_s, na.rm = TRUE),
    disttomax = mean(disttomax, na.rm = TRUE),

    rms = mean(rms, na.rm = TRUE),
    energy = mean(energy, na.rm = TRUE),
    peaktopeak = mean(peaktopeak, na.rm = TRUE),

    peak_freq_mean = mean(peak_freq_mean, na.rm = TRUE),
    peak_ampl_mean = mean(peak_ampl_mean, na.rm = TRUE),
    min_freq_mean = mean(min_freq_mean, na.rm = TRUE),
    max_freq_mean = mean(max_freq_mean, na.rm = TRUE),
    bandw_mean = mean(bandw_mean, na.rm = TRUE),

    quart_25_mean = mean(quart_25_mean, na.rm = TRUE),
    quart_50_mean = mean(quart_50_mean, na.rm = TRUE),
    quart_75_mean = mean(quart_75_mean, na.rm = TRUE),

    entropy_mean = mean(entropy_mean, na.rm = TRUE)
  ) %>%
  ungroup()
call_avgs

names(call_avgs)



```

## Squirrel level averages

Taking the averages of all the calls per animal.

```{r}


sq_avgs <- call_avgs %>%
  group_by(sq_ID, site) %>%
  summarise(
    n_calls = n(),

    # averages of call-level traits
    call_duration = mean(call_duration, na.rm = TRUE),
    no_notes = mean(no_notes, na.rm = TRUE),
    note_rate = mean(note_rate, na.rm = TRUE),

    duration = mean(duration, na.rm = TRUE),
    break_s = mean(break_s, na.rm = TRUE),
    disttomax = mean(disttomax, na.rm = TRUE),

    rms = mean(rms, na.rm = TRUE),
    energy = mean(energy, na.rm = TRUE),
    peaktopeak = mean(peaktopeak, na.rm = TRUE),

    peak_freq_mean = mean(peak_freq_mean, na.rm = TRUE),
    peak_ampl_mean = mean(peak_ampl_mean, na.rm = TRUE),
    min_freq_mean = mean(min_freq_mean, na.rm = TRUE),
    max_freq_mean = mean(max_freq_mean, na.rm = TRUE),
    bandw_mean = mean(bandw_mean, na.rm = TRUE),

    quart_25_mean = mean(quart_25_mean, na.rm = TRUE),
    quart_50_mean = mean(quart_50_mean, na.rm = TRUE),
    quart_75_mean = mean(quart_75_mean, na.rm = TRUE),

    entropy_mean = mean(entropy_mean, na.rm = TRUE)
  ) %>%
  ungroup()
sq_avgs
names(sq_avgs)



```

## 

```{r}


```

## Visualization for averages- MG

Create figures for a clear look into data.

```{r}

# Just going to look at figs for mt graham squirrels.

# using site, keep only Mount Graham call averages
call_avgs_mg <- call_avgs %>%
  filter(site == "MG")



# for MG squirrels, calculate the mean and SD of call_duration across calls
call_duration_by_sq_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_call_duration = mean(call_duration, na.rm = TRUE),
    sd_call_duration = sd(call_duration, na.rm = TRUE)
  ) %>%
  ungroup()

p_call_duration_mg <- call_duration_by_sq_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_call_duration), y = mean_call_duration)) +
  geom_pointrange(
    aes(ymin = mean_call_duration - sd_call_duration,
        ymax = mean_call_duration + sd_call_duration),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Call duration: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_call_duration_mg

ggsave("figures/var_call_duration_MG.png",
       plot = p_call_duration_mg,
       width = 8, height = 5, dpi = 300)

#  for MG squirrels, calculate mean and SD of note_rate across calls
note_rate_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_note_rate = mean(note_rate, na.rm = TRUE),
    sd_note_rate = sd(note_rate, na.rm = TRUE)
  ) %>%
  ungroup()
p_note_rate_mg <- note_rate_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_note_rate), y = mean_note_rate)) +
  geom_pointrange(
    aes(ymin = mean_note_rate - sd_note_rate,
        ymax = mean_note_rate + sd_note_rate),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Note rate: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_note_rate_mg
ggsave("figures/var_note_rate_MG.png",
       plot = p_note_rate_mg,
       width = 8, height = 5, dpi = 300)

# duration param MG
# mean + SD of duration for each MG squirrel
duration_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_duration = mean(duration, na.rm = TRUE),
    sd_duration = sd(duration, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_duration_mg <- duration_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_duration), y = mean_duration)) +
  geom_pointrange(
    aes(ymin = mean_duration - sd_duration,
        ymax = mean_duration + sd_duration),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Duration: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_duration_mg

# save into figures folder
ggsave("figures/var_duration_MG.png",
       plot = p_duration_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of break_s for each MG squirrel
break_s_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_break_s = mean(break_s, na.rm = TRUE),
    sd_break_s = sd(break_s, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_break_s_mg <- break_s_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_break_s), y = mean_break_s)) +
  geom_pointrange(
    aes(ymin = mean_break_s - sd_break_s,
        ymax = mean_break_s + sd_break_s),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "break_s: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_break_s_mg

# save into figures folder
ggsave("figures/var_break_s_MG.png",
       plot = p_break_s_mg,
       width = 8, height = 5, dpi = 300)


# mean + SD of disttomax for each MG squirrel
disttomax_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_disttomax = mean(disttomax, na.rm = TRUE),
    sd_disttomax = sd(disttomax, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_disttomax_mg <- disttomax_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_disttomax), y = mean_disttomax)) +
  geom_pointrange(
    aes(ymin = mean_disttomax - sd_disttomax,
        ymax = mean_disttomax + sd_disttomax),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "disttomax: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_disttomax_mg

# save into figures folder
ggsave("figures/var_disttomax_MG.png",
       plot = p_disttomax_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of rms for each MG squirrel
rms_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_rms = mean(rms, na.rm = TRUE),
    sd_rms = sd(rms, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_rms_mg <- rms_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_rms), y = mean_rms)) +
  geom_pointrange(
    aes(ymin = mean_rms - sd_rms,
        ymax = mean_rms + sd_rms),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "RMS: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_rms_mg

# save into figures folder
ggsave("figures/var_rms_MG.png",
       plot = p_rms_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of energy for each MG squirrel
energy_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_energy = mean(energy, na.rm = TRUE),
    sd_energy = sd(energy, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_energy_mg <- energy_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_energy), y = mean_energy)) +
  geom_pointrange(
    aes(ymin = mean_energy - sd_energy,
        ymax = mean_energy + sd_energy),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Energy: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_energy_mg

# save into figures folder
ggsave("figures/var_energy_MG.png",
       plot = p_energy_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of peaktopeak for each MG squirrel
peaktopeak_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_peaktopeak = mean(peaktopeak, na.rm = TRUE),
    sd_peaktopeak = sd(peaktopeak, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_peaktopeak_mg <- peaktopeak_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_peaktopeak), y = mean_peaktopeak)) +
  geom_pointrange(
    aes(ymin = mean_peaktopeak - sd_peaktopeak,
        ymax = mean_peaktopeak + sd_peaktopeak),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Peak-to-peak: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_peaktopeak_mg

# save into figures folder
ggsave("figures/var_peaktopeak_MG.png",
       plot = p_peaktopeak_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of peak_freq_mean for each MG squirrel
peak_freq_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_peak_freq = mean(peak_freq_mean, na.rm = TRUE),
    sd_peak_freq = sd(peak_freq_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_peak_freq_mg <- peak_freq_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_peak_freq), y = mean_peak_freq)) +
  geom_pointrange(
    aes(ymin = mean_peak_freq - sd_peak_freq,
        ymax = mean_peak_freq + sd_peak_freq),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Peak frequency mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_peak_freq_mg

# save into figures folder
ggsave("figures/var_peak_freq_mean_MG.png",
       plot = p_peak_freq_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of peak_ampl_mean for each MG squirrel
peak_ampl_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_peak_ampl = mean(peak_ampl_mean, na.rm = TRUE),
    sd_peak_ampl = sd(peak_ampl_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_peak_ampl_mg <- peak_ampl_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_peak_ampl), y = mean_peak_ampl)) +
  geom_pointrange(
    aes(ymin = mean_peak_ampl - sd_peak_ampl,
        ymax = mean_peak_ampl + sd_peak_ampl),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Peak amplitude mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_peak_ampl_mg

# save into figures folder
ggsave("figures/var_peak_ampl_mean_MG.png",
       plot = p_peak_ampl_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of min_freq_mean for each MG squirrel
min_freq_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_min_freq = mean(min_freq_mean, na.rm = TRUE),
    sd_min_freq = sd(min_freq_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_min_freq_mg <- min_freq_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_min_freq), y = mean_min_freq)) +
  geom_pointrange(
    aes(ymin = mean_min_freq - sd_min_freq,
        ymax = mean_min_freq + sd_min_freq),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Min frequency mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_min_freq_mg

# save into figures folder
ggsave("figures/var_min_freq_mean_MG.png",
       plot = p_min_freq_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of max_freq_mean for each MG squirrel
max_freq_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_max_freq = mean(max_freq_mean, na.rm = TRUE),
    sd_max_freq = sd(max_freq_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_max_freq_mg <- max_freq_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_max_freq), y = mean_max_freq)) +
  geom_pointrange(
    aes(ymin = mean_max_freq - sd_max_freq,
        ymax = mean_max_freq + sd_max_freq),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Max frequency mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_max_freq_mg

# save into figures folder
ggsave("figures/var_max_freq_mean_MG.png",
       plot = p_max_freq_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of bandw_mean for each MG squirrel
bandw_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_bandw = mean(bandw_mean, na.rm = TRUE),
    sd_bandw = sd(bandw_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_bandw_mg <- bandw_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_bandw), y = mean_bandw)) +
  geom_pointrange(
    aes(ymin = mean_bandw - sd_bandw,
        ymax = mean_bandw + sd_bandw),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Bandwidth mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_bandw_mg

# save into figures folder
ggsave("figures/var_bandw_mean_MG.png",
       plot = p_bandw_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of quart_25_mean for each MG squirrel
quart25_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_quart25 = mean(quart_25_mean, na.rm = TRUE),
    sd_quart25 = sd(quart_25_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_quart25_mg <- quart25_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_quart25), y = mean_quart25)) +
  geom_pointrange(
    aes(ymin = mean_quart25 - sd_quart25,
        ymax = mean_quart25 + sd_quart25),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Quartile 25% mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_quart25_mg

# save into figures folder
ggsave("figures/var_quart_25_mean_MG.png",
       plot = p_quart25_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of quart_50_mean for each MG squirrel
quart50_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_quart50 = mean(quart_50_mean, na.rm = TRUE),
    sd_quart50 = sd(quart_50_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_quart50_mg <- quart50_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_quart50), y = mean_quart50)) +
  geom_pointrange(
    aes(ymin = mean_quart50 - sd_quart50,
        ymax = mean_quart50 + sd_quart50),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Quartile 50% mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_quart50_mg

# save into figures folder
ggsave("figures/var_quart_50_mean_MG.png",
       plot = p_quart50_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of quart_75_mean for each MG squirrel
quart75_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_quart75 = mean(quart_75_mean, na.rm = TRUE),
    sd_quart75 = sd(quart_75_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_quart75_mg <- quart75_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_quart75), y = mean_quart75)) +
  geom_pointrange(
    aes(ymin = mean_quart75 - sd_quart75,
        ymax = mean_quart75 + sd_quart75),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Quartile 75% mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_quart75_mg

# save into figures folder
ggsave("figures/var_quart_75_mean_MG.png",
       plot = p_quart75_mg,
       width = 8, height = 5, dpi = 300)

# mean + SD of entropy_mean for each MG squirrel
entropy_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_entropy = mean(entropy_mean, na.rm = TRUE),
    sd_entropy = sd(entropy_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_entropy_mg <- entropy_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_entropy), y = mean_entropy)) +
  geom_pointrange(
    aes(ymin = mean_entropy - sd_entropy,
        ymax = mean_entropy + sd_entropy),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Entropy mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_entropy_mg

# save into figures folder
ggsave("figures/var_entropy_mean_MG.png",
       plot = p_entropy_mg,
       width = 8, height = 5, dpi = 300)


# mean + SD of note_rate for each MG squirrel (across calls)
note_rate_mg <- call_avgs_mg %>%
  group_by(sq_ID) %>%
  summarise(
    mean_note_rate = mean(note_rate, na.rm = TRUE),
    sd_note_rate = sd(note_rate, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_note_rate_mg <- note_rate_mg %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_note_rate), y = mean_note_rate)) +
  geom_pointrange(
    aes(ymin = mean_note_rate - sd_note_rate,
        ymax = mean_note_rate + sd_note_rate),
    na.rm = TRUE
  ) +
  labs(x = "Mount Graham squirrels (Squirrel ID)", y = "Note rate: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_note_rate_mg

# save it
ggsave("figures/var_note_rate_MG.png",
       plot = p_note_rate_mg,
       width = 8, height = 5, dpi = 300)





```

## Visualization for averages-WM

```{r}

call_avgs_wm <- call_avgs %>%
  filter(site == "WM")


# mean + SD (across calls) for each WM squirrel
call_duration_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_call_duration = mean(call_duration, na.rm = TRUE),
    sd_call_duration = sd(call_duration, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_call_duration_wm <- call_duration_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_call_duration), y = mean_call_duration)) +
  geom_pointrange(
    aes(ymin = mean_call_duration - sd_call_duration,
        ymax = mean_call_duration + sd_call_duration),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Call duration: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_call_duration_wm

ggsave("figures/var_call_duration_WM.png",
       plot = p_call_duration_wm,
       width = 8, height = 5, dpi = 300)

# note rate
note_rate_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_note_rate = mean(note_rate, na.rm = TRUE),
    sd_note_rate = sd(note_rate, na.rm = TRUE)
  ) %>%
  ungroup()

p_note_rate_wm <- note_rate_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_note_rate), y = mean_note_rate)) +
  geom_pointrange(
    aes(ymin = mean_note_rate - sd_note_rate,
        ymax = mean_note_rate + sd_note_rate),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Note rate: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_note_rate_wm

ggsave("figures/var_note_rate_WM.png",
       plot = p_note_rate_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of duration for each WM squirrel
duration_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_duration = mean(duration, na.rm = TRUE),
    sd_duration = sd(duration, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_duration_wm <- duration_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_duration), y = mean_duration)) +
  geom_pointrange(
    aes(ymin = mean_duration - sd_duration,
        ymax = mean_duration + sd_duration),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Duration: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_duration_wm

# save it
ggsave("figures/var_duration_WM.png",
       plot = p_duration_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of break_s for each WM squirrel
break_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_break = mean(break_s, na.rm = TRUE),
    sd_break = sd(break_s, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_break_wm <- break_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_break), y = mean_break)) +
  geom_pointrange(
    aes(ymin = mean_break - sd_break,
        ymax = mean_break + sd_break),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Break_s: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_break_wm

# save it
ggsave("figures/var_break_s_WM.png",
       plot = p_break_wm,
       width = 8, height = 5, dpi = 300)


# mean + SD (across calls) of rms for each WM squirrel
rms_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_rms = mean(rms, na.rm = TRUE),
    sd_rms = sd(rms, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_rms_wm <- rms_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_rms), y = mean_rms)) +
  geom_pointrange(
    aes(ymin = mean_rms - sd_rms,
        ymax = mean_rms + sd_rms),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "RMS: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_rms_wm

# save it
ggsave("figures/var_rms_WM.png",
       plot = p_rms_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of energy for each WM squirrel
energy_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_energy = mean(energy, na.rm = TRUE),
    sd_energy = sd(energy, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_energy_wm <- energy_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_energy), y = mean_energy)) +
  geom_pointrange(
    aes(ymin = mean_energy - sd_energy,
        ymax = mean_energy + sd_energy),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Energy: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_energy_wm

# save it
ggsave("figures/var_energy_WM.png",
       plot = p_energy_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of peaktopeak for each WM squirrel
peaktopeak_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_peaktopeak = mean(peaktopeak, na.rm = TRUE),
    sd_peaktopeak = sd(peaktopeak, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_peaktopeak_wm <- peaktopeak_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_peaktopeak), y = mean_peaktopeak)) +
  geom_pointrange(
    aes(ymin = mean_peaktopeak - sd_peaktopeak,
        ymax = mean_peaktopeak + sd_peaktopeak),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Peak-to-peak: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_peaktopeak_wm

# save it
ggsave("figures/var_peaktopeak_WM.png",
       plot = p_peaktopeak_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of peak_freq_mean for each WM squirrel
peak_freq_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_peak_freq = mean(peak_freq_mean, na.rm = TRUE),
    sd_peak_freq = sd(peak_freq_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_peak_freq_wm <- peak_freq_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_peak_freq), y = mean_peak_freq)) +
  geom_pointrange(
    aes(ymin = mean_peak_freq - sd_peak_freq,
        ymax = mean_peak_freq + sd_peak_freq),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Peak frequency mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_peak_freq_wm

# save it
ggsave("figures/var_peak_freq_mean_WM.png",
       plot = p_peak_freq_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of peak_ampl_mean for each WM squirrel
peak_ampl_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_peak_ampl = mean(peak_ampl_mean, na.rm = TRUE),
    sd_peak_ampl = sd(peak_ampl_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_peak_ampl_wm <- peak_ampl_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_peak_ampl), y = mean_peak_ampl)) +
  geom_pointrange(
    aes(ymin = mean_peak_ampl - sd_peak_ampl,
        ymax = mean_peak_ampl + sd_peak_ampl),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Peak amplitude mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_peak_ampl_wm

# save it
ggsave("figures/var_peak_ampl_mean_WM.png",
       plot = p_peak_ampl_wm,
       width = 8, height = 5, dpi = 300)


# mean + SD (across calls) of min_freq_mean for each WM squirrel
min_freq_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_min_freq = mean(min_freq_mean, na.rm = TRUE),
    sd_min_freq = sd(min_freq_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_min_freq_wm <- min_freq_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_min_freq), y = mean_min_freq)) +
  geom_pointrange(
    aes(ymin = mean_min_freq - sd_min_freq,
        ymax = mean_min_freq + sd_min_freq),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Min frequency mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_min_freq_wm

# save it
ggsave("figures/var_min_freq_mean_WM.png",
       plot = p_min_freq_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of max_freq_mean for each WM squirrel
max_freq_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_max_freq = mean(max_freq_mean, na.rm = TRUE),
    sd_max_freq = sd(max_freq_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_max_freq_wm <- max_freq_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_max_freq), y = mean_max_freq)) +
  geom_pointrange(
    aes(ymin = mean_max_freq - sd_max_freq,
        ymax = mean_max_freq + sd_max_freq),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Max frequency mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_max_freq_wm

# save it
ggsave("figures/var_max_freq_mean_WM.png",
       plot = p_max_freq_wm,
       width = 8, height = 5, dpi = 300)


# mean + SD (across calls) of bandw_mean for each WM squirrel
bandw_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_bandw = mean(bandw_mean, na.rm = TRUE),
    sd_bandw = sd(bandw_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_bandw_wm <- bandw_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_bandw), y = mean_bandw)) +
  geom_pointrange(
    aes(ymin = mean_bandw - sd_bandw,
        ymax = mean_bandw + sd_bandw),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Bandwidth mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_bandw_wm

# save it
ggsave("figures/var_bandw_mean_WM.png",
       plot = p_bandw_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of quart_25_mean for each WM squirrel
quart25_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_quart25 = mean(quart_25_mean, na.rm = TRUE),
    sd_quart25 = sd(quart_25_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_quart25_wm <- quart25_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_quart25), y = mean_quart25)) +
  geom_pointrange(
    aes(ymin = mean_quart25 - sd_quart25,
        ymax = mean_quart25 + sd_quart25),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Quartile 25% mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_quart25_wm

# save it
ggsave("figures/var_quart_25_mean_WM.png",
       plot = p_quart25_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of quart_50_mean for each WM squirrel
quart50_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_quart50 = mean(quart_50_mean, na.rm = TRUE),
    sd_quart50 = sd(quart_50_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_quart50_wm <- quart50_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_quart50), y = mean_quart50)) +
  geom_pointrange(
    aes(ymin = mean_quart50 - sd_quart50,
        ymax = mean_quart50 + sd_quart50),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Quartile 50% mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_quart50_wm

# save it
ggsave("figures/var_quart_50_mean_WM.png",
       plot = p_quart50_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of quart_75_mean for each WM squirrel
quart75_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_quart75 = mean(quart_75_mean, na.rm = TRUE),
    sd_quart75 = sd(quart_75_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_quart75_wm <- quart75_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_quart75), y = mean_quart75)) +
  geom_pointrange(
    aes(ymin = mean_quart75 - sd_quart75,
        ymax = mean_quart75 + sd_quart75),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Quartile 75% mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_quart75_wm

# save it
ggsave("figures/var_quart_75_mean_WM.png",
       plot = p_quart75_wm,
       width = 8, height = 5, dpi = 300)

# mean + SD (across calls) of entropy_mean for each WM squirrel
entropy_wm <- call_avgs_wm %>%
  group_by(sq_ID) %>%
  summarise(
    mean_entropy = mean(entropy_mean, na.rm = TRUE),
    sd_entropy = sd(entropy_mean, na.rm = TRUE)
  ) %>%
  ungroup()

# plot mean ± SD
p_entropy_wm <- entropy_wm %>%
  ggplot(aes(x = reorder(factor(sq_ID), mean_entropy), y = mean_entropy)) +
  geom_pointrange(
    aes(ymin = mean_entropy - sd_entropy,
        ymax = mean_entropy + sd_entropy),
    na.rm = TRUE
  ) +
  labs(x = "White Mountains squirrels (Squirrel ID)", y = "Entropy mean: mean ± SD (across calls)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

p_entropy_wm

# save it
ggsave("figures/var_entropy_mean_WM.png",
       plot = p_entropy_wm,
       width = 8, height = 5, dpi = 300)


```

## ICC MG

Do mt graham red squirrels have individual distinctiveness in their calls? Do White Mountain squirrels have distiveness in their calls? I will run a one way ANOVA and then control for inflated p-values by running a FDR/BH sensitivity test.

I will also run a mixed model to better explain variation.

```{r}
########### Run a mixed model ###############

# Mount Graham call-level data for mixed models
mg_df <- call_avgs_mg



# Just doing one parameter to test 'entropy_mean'
mg_entropy_mod <- lmer(entropy_mean ~ 1 + (1 | sq_ID), data = mg_df)
summary(mg_entropy_mod)

# compute ICC

mg_entropy_vc <- VarCorr(mg_entropy_mod)

mg_var_sq_ID <- as.numeric(mg_entropy_vc$sq_ID[1])
mg_var_resid  <- attr(mg_entropy_vc, "sc")^2

mg_icc_entropy <- mg_var_sq_ID / (mg_var_sq_ID + mg_var_resid)
mg_icc_entropy

# grand means

mg_entropy_effects <- ranef(mg_entropy_mod)$sq_ID[, 1]
mg_entropy_grand_mean <- fixef(mg_entropy_mod)[1]

mg_entropy_means <- mg_entropy_grand_mean + mg_entropy_effects
mg_entropy_means


# Everything looks good, so I will run all parameters


mg_model_results <- lapply(params, function(param_name) {

  # fit model for one parameter
  mg_mod <- lmer(
    as.formula(paste0(param_name, " ~ 1 + (1 | sq_ID)")),
    data = mg_df
  )

  # variance components -> ICC
  vc <- VarCorr(mg_mod)
  var_sq_ID <- as.numeric(vc$sq_ID[1])
  var_resid  <- attr(vc, "sc")^2
  icc_val <- var_sq_ID / (var_sq_ID + var_resid)

  # return a small list
  list(
    parameter = param_name,
    model = mg_mod,
    var_sq_ID = var_sq_ID,
    var_resid = var_resid,
    icc = icc_val
  )
})
mg_icc_results <- tibble(
  parameter  = sapply(mg_model_results, `[[`, "parameter"),
  var_sq_ID = sapply(mg_model_results, `[[`, "var_sq_ID"),
  var_resid  = sapply(mg_model_results, `[[`, "var_resid"),
  icc        = sapply(mg_model_results, `[[`, "icc")
) %>%
  arrange(desc(icc))

mg_icc_results

write.csv(mg_icc_results, "tables/mg_icc_results.csv", row.names = FALSE)

# plot results
mg_icc_results %>%
  arrange(icc) %>%
  ggplot(aes(x = icc, y = reorder(parameter, icc))) +
  geom_point(na.rm = TRUE) +
  labs(
    x = "ICC (proportion of variance among squirrels)",
    y = "Parameter",
    title = "Mount Graham: individual distinctiveness by parameter (ICC)"
  )


p_mg_icc <- mg_icc_results %>%
  arrange(icc) %>%
  ggplot(aes(x = icc, y = reorder(parameter, icc))) +
  geom_point(na.rm = TRUE) +
  labs(
    x = "ICC (proportion of variance among squirrels)",
    y = "Parameter",
    title = "Mount Graham: ICC for each acoustic parameter"
  )

p_mg_icc
ggsave("figures/mg_icc_dotplot.png", plot = p_mg_icc, width = 8, height = 5, dpi = 300)

```

## ICC WM

Run a mixed model for the white mountain squirrels

```{r}
# White Mountains: mixed models + ICC table + dotplot

#  WM call-level data
wm_df <- call_avgs_wm

#  Run all WM models + ICC
wm_model_results <- lapply(params, function(param_name) {

  wm_mod <- lmer(
    as.formula(paste0(param_name, " ~ 1 + (1 | sq_ID)")),
    data = wm_df
  )

  vc <- VarCorr(wm_mod)
  var_sq_ID <- as.numeric(vc$sq_ID[1])
  var_resid  <- attr(vc, "sc")^2
  icc_val <- var_sq_ID / (var_sq_ID + var_resid)

  list(
    parameter = param_name,
    model = wm_mod,
    var_sq_ID = var_sq_ID,
    var_resid = var_resid,
    icc = icc_val
  )
})

#  WM ICC results table
wm_icc_results <- tibble(
  parameter  = sapply(wm_model_results, `[[`, "parameter"),
  var_sq_ID = sapply(wm_model_results, `[[`, "var_sq_ID"),
  var_resid  = sapply(wm_model_results, `[[`, "var_resid"),
  icc        = sapply(wm_model_results, `[[`, "icc")
) %>%
  arrange(desc(icc))

wm_icc_results
write.csv(wm_icc_results, "tables/wm_icc_results.csv", row.names = FALSE)

# WM ICC dot plot
p_wm_icc <- wm_icc_results %>%
  arrange(icc) %>%
  ggplot(aes(x = icc, y = reorder(parameter, icc))) +
  geom_point(na.rm = TRUE) +
  labs(
    x = "ICC (proportion of variance among squirrels)",
    y = "Parameter",
    title = "White Mountains: individual distinctiveness by parameter (ICC)"
  )

p_wm_icc
ggsave("figures/wm_icc_dotplot.png", plot = p_wm_icc, width = 8, height = 5, dpi = 300)

```

```{r}
# add site labels to each ICC table
wm_icc_results <- wm_icc_results %>%
  mutate(site = "White Mountains")

mg_icc_results <- mg_icc_results %>%
  mutate(site = "Mount Graham")

# combine for plotting
icc_results_both <- bind_rows(wm_icc_results, mg_icc_results)

# facet plot: keeps site-specific ordering without "combining" in reorder()
p_icc_facet <- icc_results_both %>%
  mutate(param_site = paste(parameter, site, sep = "___")) %>%
  ggplot(aes(x = icc, y = reorder(param_site, icc))) +
  geom_point(size = 3, na.rm = TRUE) +
  facet_wrap(~ site, scales = "free_y") +
  scale_y_discrete(labels = function(x) sub("___.*$", "", x)) +
  labs(
    x = "ICC (proportion of variance among squirrels)",
    y = "Parameter",
    title = "Individual distinctiveness by parameter (ICC)"
  ) +
  theme_classic() +
  theme(
    plot.title  = element_text(size = 18, face = "bold"),
    strip.text  = element_text(size = 16, face = "bold"),
    axis.title.x = element_text(size = 15),
    axis.title.y = element_text(size = 15)
  )

p_icc_facet


ggsave(
  filename = "figures/icc_facet_sites.png",
  plot = p_icc_facet,
  width = 11,
  height = 7,
  dpi = 300
)


```

## Sq variation catepillar plot

```{r}

```

## Stats-Grand mean/ SD and PICs

tt \<- gridExtra::ttheme_default(base_size = 10)

pic_mg \<- pic_table_2 %\>%

filter(site == "Mount Graham") %\>%

select(parameter, CVa, mean_CVw, PIC, eta2) %\>%

rename(\`Call property\` = parameter)

pic_wm \<- pic_table_2 %\>%

filter(site == "White Mountains") %\>%

select(parameter, CVa, mean_CVw, PIC, eta2) %\>%

rename(\`Call property\` = parameter)

png("figures/pic_table_mg.png", width = 1200, height = 700, res = 200)

gridExtra::grid.table(pic_mg, theme = tt); dev.off()

png("figures/pic_table_wm.png", width = 1200, height = 700, res = 200)

gridExtra::grid.table(pic_wm, theme = tt); dev.off()

```{r}

# Output object: PIC summary table
pic_summary_by_site <- data.frame()

for (current_site in unique(call_avgs$site)) {

  call_avgs_site <- call_avgs %>%
    filter(site == current_site)

  for (param in params) {

    # Within-individual stats (across calls)
    ind_stats <- call_avgs_site %>%
      group_by(sq_ID) %>%
      summarise(
        ind_mean = mean(.data[[param]], na.rm = TRUE),
        ind_sd   = sd(.data[[param]], na.rm = TRUE),
        .groups = "drop"
      ) %>%
      mutate(
        CVw = (ind_sd / ind_mean) * 100
      )

    # Among-individual (grand) stats
    grand_mean <- mean(ind_stats$ind_mean, na.rm = TRUE)
    grand_sd   <- sd(ind_stats$ind_mean, na.rm = TRUE)
    CVa        <- (grand_sd / grand_mean) * 100

    mean_CVw <- mean(ind_stats$CVw, na.rm = TRUE)
    min_CVw  <- min(ind_stats$CVw, na.rm = TRUE)
    max_CVw  <- max(ind_stats$CVw, na.rm = TRUE)

    PIC <- CVa / mean_CVw

    # eta^2 (R^2 from one-way ANOVA)
    model_data <- call_avgs_site %>%
      select(sq_ID, all_of(param)) %>%
      filter(!is.na(.data[[param]]))

    eta2 <- summary(lm(aov(model_data[[param]] ~ as.factor(model_data$sq_ID))))$r.squared

    # Append one row to summary table
    pic_summary_by_site <- rbind(
      pic_summary_by_site,
      data.frame(
        site       = current_site,
        parameter  = param,
        grand_mean = grand_mean,
        grand_sd   = grand_sd,
        CVa        = CVa,
        mean_CVw   = mean_CVw,
        min_CVw    = min_CVw,
        max_CVw    = max_CVw,
        PIC        = PIC,
        eta2       = eta2,
        n_IDs      = n_distinct(model_data$sq_ID),
        n_calls    = nrow(model_data)
      )
    )
  }
}

pic_summary_by_site <- as_tibble(pic_summary_by_site)

# dropiing the number of calls and sq IDs from the table
pic_table_1 <- pic_summary_by_site %>%
  select(-n_IDs, -n_calls)

# cleaning up decimals
pic_table_2 <- pic_table_1 %>%
  mutate(
    grand_mean = round(grand_mean, 2),
    grand_sd   = round(grand_sd, 2),
    CVa        = round(CVa, 2),
    mean_CVw   = round(mean_CVw, 2),
    min_CVw    = round(min_CVw, 2),
    max_CVw    = round(max_CVw, 2),
    PIC        = round(PIC, 2),
    eta2       = round(eta2, 2),

    # relabel sites (WM/MG -> full names)
    site = case_when(
      site == "WM" ~ "White Mountains",
      site == "MG" ~ "Mount Graham",
      TRUE ~ site
    )
  )

pic_table_2

write_csv(pic_table_2, "tables/pic_table_PIC_combined.csv")


library(gridExtra)
library(grid)
library(gridExtra)

# bigger table text
tt <- gridExtra::ttheme_default(base_size = 22)

pic_mg <- pic_table_2 %>%
  filter(site == "Mount Graham") %>%
  select(parameter, CVa, mean_CVw, PIC, eta2) %>%
  mutate(parameter = gsub("_", " ", parameter)) %>%
  rename(`Call property` = parameter)

title_mg <- grid::textGrob(
  "Mount Graham Red Squirrels: PIC Summary by Acoustic Parameter",
  gp = grid::gpar(fontsize = 19, fontface = "bold")
)

table_mg <- gridExtra::tableGrob(pic_mg, rows = NULL, theme = tt)

pdf("figures/pic_table_mount_graham.pdf", width = 8.5, height = 11)
grid::grid.newpage()
grid::grid.draw(
  gridExtra::arrangeGrob(
    title_mg, table_mg,
    ncol = 1,
    heights = c(0.3, 10.7)   # smaller title row = less whitespace
  )
)
dev.off()



tt <- gridExtra::ttheme_default(base_size = 22)

pic_wm <- pic_table_2 %>%
  filter(site == "White Mountains") %>%
  select(parameter, CVa, mean_CVw, PIC, eta2) %>%
  mutate(parameter = gsub("_", " ", parameter)) %>%
  rename(`Call property` = parameter)

title_wm <- grid::textGrob(
  "White Mountains Red Squirrels: PIC Summary by Acoustic Parameter",
  gp = grid::gpar(fontsize = 16, fontface = "bold")
)

table_wm <- gridExtra::tableGrob(pic_wm, rows = NULL, theme = tt)

pdf("figures/pic_table_white_mountains.pdf", width = 8.5, height = 11)
grid::grid.newpage()
grid::grid.draw(
  gridExtra::arrangeGrob(
    title_wm, table_wm,
    ncol = 1,
    heights = c(0.3, 10.7)
  )
)
dev.off()

```

## Hs (Beechers stat)

Check to see if there is any recording bias by distance. Plot distance = x and parameter = y for every parameter for each population. Then compute Hs beechers stat.

```{r}
# Hs (Beecher's information statistic)
# calculate Hs (in bits) for each acoustic parameter, separately for MG and WM

# function: compute Beecher's Hs for ONE parameter in MG calls
calc_Hs_one_parameter_MG <- function(parameter) {

  hs_model_MG <- lme4::lmer(
    stats::as.formula(paste0(parameter, " ~ 1 + (1 | sq_ID)")),
    data = call_avgs_mg,
    REML = TRUE
  )

  hs_variance_MG <- as.data.frame(lme4::VarCorr(hs_model_MG))

  var_among_squirrels <- hs_variance_MG$vcov[hs_variance_MG$grp == "sq_ID"]
  var_within_squirrels <- hs_variance_MG$vcov[hs_variance_MG$grp == "Residual"]

  Hs <- log2((var_among_squirrels + var_within_squirrels) / var_within_squirrels)

  tibble::tibble(
    site = "MG",
    parameter = parameter,
    var_among_squirrels = var_among_squirrels,
    var_within_squirrels = var_within_squirrels,
    Hs = Hs
  )
}

# function: compute Beecher's Hs for ONE parameter in WM calls
calc_Hs_one_parameter_WM <- function(parameter) {

  hs_model_WM <- lme4::lmer(
    stats::as.formula(paste0(parameter, " ~ 1 + (1 | sq_ID)")),
    data = call_avgs_wm,
    REML = TRUE
  )

  hs_variance_WM <- as.data.frame(lme4::VarCorr(hs_model_WM))

  var_among_squirrels <- hs_variance_WM$vcov[hs_variance_WM$grp == "sq_ID"]
  var_within_squirrels <- hs_variance_WM$vcov[hs_variance_WM$grp == "Residual"]

  Hs <- log2((var_among_squirrels + var_within_squirrels) / var_within_squirrels)

  tibble::tibble(
    site = "WM",
    parameter = parameter,
    var_among_squirrels = var_among_squirrels,
    var_within_squirrels = var_within_squirrels,
    Hs = Hs
  )
}

# run Hs for all parameters (MG)
Hs_table_MG <- purrr::map_dfr(params, calc_Hs_one_parameter_MG)

# run Hs for all parameters (WM)
Hs_table_WM <- purrr::map_dfr(params, calc_Hs_one_parameter_WM)

# combine MG + WM into one table
Hs_table <- bind_rows(Hs_table_MG, Hs_table_WM)

# round 2 decimals
Hs_table_clean <- Hs_table %>%
  mutate(
    Hs = round(Hs, 2),
    var_among_squirrels = round(var_among_squirrels, 6),
    var_within_squirrels = round(var_within_squirrels, 6)
  )

# view final table
Hs_table_clean
write_csv(Hs_table_clean, "tables/Hs_table_clean.csv")


# Create a more presentable table splitting each population up.



tt <- gridExtra::ttheme_default(base_size = 20)

# make a readable version (full site names + nicer parameter text)
Hs_table_clean_2 <- Hs_table_clean %>%
  mutate(
    site = case_when(
      site == "MG" ~ "Mount Graham",
      site == "WM" ~ "White Mountains",
      TRUE ~ site
    ),
    parameter = gsub("_", " ", parameter)
  )

# --- Mount Graham Hs table ---
Hs_mg <- Hs_table_clean_2 %>%
  filter(site == "Mount Graham") %>%
  select(parameter, Hs, var_among_squirrels, var_within_squirrels) %>%
  rename(`Call property` = parameter,
         `Var among squirrels` = var_among_squirrels,
         `Var within squirrels` = var_within_squirrels)

title_hs_mg <- grid::textGrob(
  "Mount Graham Red Squirrels: Beecher's Hs by Acoustic Parameter",
  gp = grid::gpar(fontsize = 16, fontface = "bold")
)

table_hs_mg <- gridExtra::tableGrob(Hs_mg, rows = NULL, theme = tt)

pdf("figures/hs_table_mount_graham.pdf", width = 8.5, height = 11)
grid::grid.newpage()
grid::grid.draw(gridExtra::arrangeGrob(title_hs_mg, table_hs_mg, ncol = 1, heights = c(0.3, 10.7)))
dev.off()

# --- White Mountains Hs table ---
Hs_wm <- Hs_table_clean_2 %>%
  filter(site == "White Mountains") %>%
  select(parameter, Hs, var_among_squirrels, var_within_squirrels) %>%
  rename(`Call property` = parameter,
         `Var among squirrels` = var_among_squirrels,
         `Var within squirrels` = var_within_squirrels)

title_hs_wm <- grid::textGrob(
  "White Mountains Red Squirrels: Beecher's Hs by Acoustic Parameter",
  gp = grid::gpar(fontsize = 16, fontface = "bold")
)

table_hs_wm <- gridExtra::tableGrob(Hs_wm, rows = NULL, theme = tt)

pdf("figures/hs_table_white_mountains.pdf", width = 8.5, height = 11)
grid::grid.newpage()
grid::grid.draw(gridExtra::arrangeGrob(title_hs_wm, table_hs_wm, ncol = 1, heights = c(0.3, 10.7)))
dev.off()

# make a comparizon figure that shows comparison between wm Hs and mg Hs
# Bar graph. MG vs WM side-by-side for each parameter
p_hs_bar <- Hs_table_clean_2 %>%
  ggplot(aes(x = parameter, y = Hs, fill = site)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  coord_flip() +
  scale_fill_manual(values = c("Mount Graham" = "#0072B2",  # blue
                               "White Mountains" = "#E69F00")) + # orange
  labs(
    title = "Beecher's Hs by acoustic parameter (MG vs WM)",
    x = NULL,
    y = "Hs (bits)",
    fill = "Population"
  ) +
  theme_bw() +
  theme(
    legend.position = "top",
    plot.title = element_text(face = "bold", size = 18),
    
    # axis tick labels (more readable)
    axis.text.y = element_text(size = 14),  # parameter names
    axis.text.x = element_text(size = 14),  # Hs tick numbers
    
    # axis titles
    axis.title.x = element_text(size = 16), # this is "Hs (bits)" after coord_flip()
    axis.title.y = element_text(size = 16),
    
    # optional: legend text/title sizes to match
    legend.title = element_text(size = 14),
    legend.text  = element_text(size = 13)
  )

p_hs_bar

ggsave("figures/hs_compare_bar.png", plot = p_hs_bar, width = 10, height = 7, dpi = 300)


```

## Create fake forshadowing plots

```{r}
# some of my parameters
fake_params <- c(
  "call duration", "note rate", "peak freq", "bandwidth",
  "entropy", "rms", "energy", "break"
)

# -----------------------------
# Fake dataset 1: NO individuality (low Hs, basically flat)
# -----------------------------
hs_fake_no_individuality <- tibble(
  parameter  = fake_params,
  Hs = c(0.18, 0.20, 0.17, 0.19, 0.16, 0.15, 0.14, 0.18)
) %>%
  mutate(parameter = factor(parameter, levels = fake_params))

# -----------------------------
# Fake dataset 2: HIGH individuality (much higher Hs, more dramatic)
# -----------------------------
hs_fake_high_individuality <- tibble(
  parameter  = fake_params,
  Hs = c(3.8, 3.1, 5.2, 3.6, 4.4, 2.7, 2.4, 3.0)
) %>%
  mutate(parameter = factor(parameter, levels = fake_params))

# -----------------------------
# Plot function so both match style
# -----------------------------
plot_fake_hs <- function(df, title_text) {
  ggplot(df, aes(x = parameter, y = Hs)) +
    geom_col(na.rm = TRUE) +
    coord_flip() +
    labs(
      title = title_text,
      x = NULL,
      y = "Hs (bits)"
    ) +
    theme_classic() +
    theme(
      axis.text = element_text(size = 14),     # tick labels
      axis.title = element_text(size = 16),    # axis titles
      plot.title = element_text(size = 18)     # plot title
    )
}

# Make the two plots
plot_fake_hs(hs_fake_no_individuality, "Hs: No individuality")
plot_fake_hs(hs_fake_high_individuality, "Hs: High individuality")
p_no <- plot_fake_hs(hs_fake_no_individuality, "Hs: No individuality")
p_hi <- plot_fake_hs(hs_fake_high_individuality, "Hs: High individuality")
ggsave("figures/hs_no_individuality.png", plot = p_no, width = 7, height = 4.5, dpi = 300)
ggsave("figures/hs_high_individuality.png", plot = p_hi, width = 7, height = 4.5, dpi = 300)
```
